### 2022-02-14 title-page.tex
Adds title page for new journal

% molde para cartas con membrete Dpto. y para sobres con ventana direcci¢n
\documentclass{letter}
\usepackage{graphicx}
\usepackage[spanish]{babel}      % castellano
\usepackage[latin1]{inputenc}    % tildes
\usepackage{url}

% para cartas en ingl‚s comentar la linea anterior e incluir la siguiente:
%\documentstyle[12pt]{a4letter}
% cabecera para hojas con membrete del Departamento y JJ
% Set left margin - The default is 1 inch, so the following 
% command sets a 0.5-inch left margin.
\setlength{\oddsidemargin}{0.0in}

% Set width of the text - What is left will be the right margin.
% In this case, right margin is 8.5in - 0.0in - 6.5in = 2.0in.
\setlength{\textwidth}{6.5in}

% Set top margin - The default is 1 inch, so the following 
% command sets a 0.5-inch top margin.
\setlength{\topmargin}{-1in}

% Set height of the text - What is left will be the bottom margin.
% In this case, bottom margin is 11in - 0.75in - 9.5in = 0.75in
\setlength{\textheight}{10in}

% cambiar estas definiciones para personalizar el documento
\def\yo{\sc Juan J. Merelo}
\def\email{\tt jmerelo@geneura.ugr.es}
\def\tel{34-958-243162}
\def\fax{34-958-248993}

% definiciones comunes
\def\granada{18071 Granada (Spain)}
\def\ucm{Universidad de Granada}
\def\fac{ETS Ingenierías Informática y Telecomucnicaciones}
\def\dpto{Departamento de Arquitectura y Tecnolog\'ia de Computadores}

\def\cabecera{

\begin{minipage}{2.5cm}
\centerline{  \includegraphics[width=2.5cm,height=2.5cm]{escudo.eps} }
\end{minipage}
\ \
\begin{minipage}{13cm}
 \rule{13cm}{0.6mm}
 \vspace*{0.3cm}
 {\small{\bf \dpto \hfill}} \\
 {\footnotesize{\bf \fac \hfill Fax: \fax}} \\
 {\footnotesize{\bf \ucm \hfill Tel. / Phone: \tel}} \\
 {\footnotesize{\bf \granada \hfill E-mail: \email}}\\
 \rule{13cm}{0.6mm}
\end{minipage}
}

%\input cabecera
\pagestyle{empty}

\begin{document}

\cabecera

\vspace{0.5cm}

\hspace{7cm}

\hfill Granada, \today

\vspace{0.5cm}
\ \
% **** si no se quiere recuadro en la direcci¢n, comentar la linea siguiente
%\framebox[9cm][l]
{
\begin{minipage}[t]{9cm}
To whom it may concern
% ******* Poner aqu¡ nombre y direcci¢n del destinatario
\end{minipage}
 }

\vspace{0.5cm}

% ********** comienzo del texto  *******************
Please find in this submission the manuscript for our research paper  to be
considered for publication in your esteemed journal. I have tried to address all
concerns that caused the initial manuscript to be sent back, I would ask again to
be a bit more precise if this is not accepted either. In this submission I have
eliminated all author information, including the (fake) authors that were
included in the sample file provided by the journal.

This information required in the cover letter:
\begin{itemize}
\item {\bf Title of the manuscript} {\em
  Chatbots and messaging platforms in the classroom: an analysis from the
  teacher's perspective}.
\item {Authors} Juan J. Merelo (1), Pedro A. Castillo (1)
  , Antonio
M. Mora (2) , Francisco Barranco (1) ,
Noorhan Abbas (3), Alberto Guillén (1) and Olia
Tsivitanidou (4).
\item {\bf Author Affiliationa}: Numbers refer to the digits in parentheses
  above:\begin{enumerate}
  \item Department of Computer Architecture and Technology,
  University of Granada, Spain;
\item  Departamento of Signal Theory and
  Telecommunications, University of Granada, Spain;
\item Leeds University,
  Leeds, UK,
\item  CYENS, Nicosia, Cyprus.
  \end{enumerate}
\item {\bf Corresponding author details} Corresponding authors are JJ Merelo
  (jmerelo@ugr.es) and Francisco Barranco (fbarranco@ugr.es). Please address all
  communications to us both.
\item {\bf Complete co-authors email addresses} In author order (marked with an
  asterisk, corresponding authors):\begin{itemize}
  \item Juan J. Merelo* {\tt jmerelo@ugr.es}, 
  \item Pedro A. Castillo 
  {\tt pacv@ugr.es}, 
\item Antonio
M. Mora {\tt amorag@ugr.es}, 
\item Francisco Barranco*  {\tt fbarranco@ugr.es},
\item 
Noorhan Abbas  {\tt n.h.abbas@leeds.ac.uk}, 
\item Alberto Guillén  {\tt
  aguillen@ugr.es} 
\item and Olia
  Tsivitanidou  {\tt o.tsivitanidou@cyens.org.cy}.
  \end{itemize}
\end{itemize}

\newpage

\textsc{Declarations}

\textbf{Competing interests}

The authors declare that they have no competing interests.

\textbf{Authors' contributions}

All the authors have contributed to the study presented in this manuscript.

JJM has been the leader, drafted the first versions of the survey and supervised the writing of the manuscript.

PAC has revised the paper and researched the state of the art, as well as
contributed to the survey design.

AMM has revised the paper and researched the state of the art, as well as contributed to the survey design.

FB has made the analysis and processed data and contributed to the survey design.

NA has supervised the state of the art, and written the bulk of it, as well as contributed to discussion and conclusions.

OT has performed data analysis, and contributed to discussion and conclusions.

AG has contributed to the design of the survey, analysis, discussion and conclusions.

All authors have read and approved the final manuscript.

\textbf{Availability of data and materials}

The datasets generated and analysed during the current study are available in
the {\tt edubots-paper} repository, accessible from {\tt https://github.com/JJ/edubots-paper/tree/main/data}.

\textbf{Funding}

This work has been supported by EDUBOTS project, funded under the scheme Erasmus + KA2: Cooperation for innovation and the exchange of good practices - Knowledge Alliances (grant agreement no: 612446).

\textbf{Acknowledgements}

The authors want to thank to the educators who participated in the survey, and
also people who have distributed the survey in their mailing lists, especially
Mario García Valdez, Jesús Moreno León and José Raúl Canay.

% ********** fin del texto  *******************

\hspace*{8cm} Yours,  \\

\centerline{\includegraphics[width=2.5cm,height=2.5cm]{jj_sign.eps}}

\hspace*{8cm} Juan J. Merelo\\
\hspace*{8cm} Catedrático de Universidad (Full professor)

\end{document}