### 2022-09-14 response-reviewers-2.tex
Starting to write a response to the reviewers

% molde para cartas con membrete Dpto. y para sobres con ventana direcci¢n
\documentclass{letter}
\usepackage{graphicx}
\usepackage[spanish]{babel}      % castellano
\usepackage[latin1]{inputenc}    % tildes
\usepackage{url}

% para cartas en ingls comentar la linea anterior e incluir la siguiente:
%\documentstyle[12pt]{a4letter}
% cabecera para hojas con membrete del Departamento y JJ
% Set left margin - The default is 1 inch, so the following 
% command sets a 0.5-inch left margin.
\setlength{\oddsidemargin}{0.0in}

% Set width of the text - What is left will be the right margin.
% In this case, right margin is 8.5in - 0.0in - 6.5in = 2.0in.
\setlength{\textwidth}{6.5in}

% Set top margin - The default is 1 inch, so the following 
% command sets a 0.5-inch top margin.
\setlength{\topmargin}{-1in}

% Set height of the text - What is left will be the bottom margin.
% In this case, bottom margin is 11in - 0.75in - 9.5in = 0.75in
\setlength{\textheight}{10in}

% cambiar estas definiciones para personalizar el documento
\def\yo{\sc Juan J. Merelo}
\def\email{\tt jmerelo@geneura.ugr.es}
\def\tel{34-958-243162}
\def\fax{34-958-248993}

% definiciones comunes
\def\granada{18071 Granada (Spain)}
\def\ucm{Universidad de Granada}
\def\fac{ETS Ingenierías Informática y Telecomucnicaciones}
\def\dpto{Departamento de Arquitectura y Tecnolog\'ia de Computadores}

\def\cabecera{

\begin{minipage}{2.5cm}
\centerline{  \includegraphics[width=2.5cm,height=2.5cm]{escudo.eps} }
\end{minipage}
\ \
\begin{minipage}{13cm}
 \rule{13cm}{0.6mm}
 \vspace*{0.3cm}
 {\small{\bf \dpto \hfill}} \\
 {\footnotesize{\bf \fac \hfill Fax: \fax}} \\
 {\footnotesize{\bf \ucm \hfill Tel. / Phone: \tel}} \\
 {\footnotesize{\bf \granada \hfill E-mail: \email}}\\
 \rule{13cm}{0.6mm}
\end{minipage}
}

%\input cabecera
\pagestyle{empty}

\begin{document}

\cabecera

\vspace{0.5cm}

\hspace{7cm}

\hfill Granada, \today

\vspace{0.5cm}
\ \
% **** si no se quiere recuadro en la direcci¢n, comentar la linea siguiente
%\framebox[9cm][l]
{
\begin{minipage}[t]{9cm}
To whom it may concern
% ******* Poner aqu¡ nombre y direcci¢n del destinatario
\end{minipage}
 }

\vspace{0.5cm}

% ********** comienzo del texto  *******************

{\bf Response to reviewers' comments}

We are very grateful to the reviewers for their kind comments. We have rewritten
the paper quite extensively, incorporating requested changes, but also
attempting to give another angle that makes this paper apart from what was
already published in this paper

Merelo, J.J. et al. (2022). Exploring the Role of Chatbots and Messaging
Applications in Higher Education: A Teacher’s Perspective. In: Zaphiris, P.,
Ioannou, A. (eds) Learning and Collaboration Technologies. Novel Technological
Environments. HCII 2022. Lecture Notes in Computer Science, vol 13329. Springer,
Cham. https://doi.org/10.1007/978-3-031-05675-8_16

This explains that, in some cases, conclusions have been extended or
reinterpreted over the ones published in that paper. Additionally, we validated
the survey, as was also requested by the reviewers.

Next we will address the concerns of the reviewers individually.

% ********** fin del texto  *******************

\hspace*{8cm} Yours,  \\

\centerline{\includegraphics[width=2.5cm,height=2.5cm]{jj_sign.eps}}

\hspace*{8cm} Juan J. Merelo\\
\hspace*{8cm} Catedrático de Universidad (Full professor)

\end{document}