### 2023-01-08 edubots_surveys.Rnw
rev discussion - following paragraphs

%% BioMed_Central_Tex_Template_v1.06
%%                                      %
%  bmc_article.tex            ver: 1.06 %
%                                       %

%%IMPORTANT: do not delete the first line of this template
%%It must be present to enable the BMC Submission system to
%%recognize this template!!

\documentclass{bmcart}

%%% Load packages
\usepackage{amsthm,amsmath}
\usepackage{multirow}
\usepackage{rotating}
\usepackage[utf8]{inputenc} %unicode support

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                             %%
%%  If you wish to display your graphics for   %%
%%  your own use using includegraphic or       %%
%%  includegraphics, then comment out the      %%
%%  following two lines of code.               %%
%%  NB: These line *must* be included when     %%
%%  submitting to BMC.                         %%
%%  All figure files must be submitted as      %%
%%  separate graphics through the BMC          %%
%%  submission process, not included in the    %%
%%  submitted article.                         %%
%%                                             %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\def\includegraphic{}
%\def\includegraphics{}

%%% Put your definitions there:
\startlocaldefs
\endlocaldefs

%%% Begin ...
\begin{document}

%%% Start of article front matter
\begin{frontmatter}

\begin{fmbox}
\dochead{Research}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Enter the title of your article here     %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title{Chatbots and messaging platforms in the classroom: an analysis from the teacher's perspective}

% Remember to substitute figures/ by nothing in the generated PDF
% \author[
%   addressref={aff1},
%   corref={aff1},
%   email={jmerelo@ugr.es}
% ]{\inits{JJM}\fnm{Juan J.} \snm{Merelo}}
% \author[
%   addressref={aff1},
%   email={pacv@ugr.es}
% ]{\inits{PAC}\fnm{Pedro A.} \snm{Castillo}}
% \author[
%   addressref={aff2},
%   email={amorag@ugr.es}
% ]{\inits{AMM}\fnm{Antonio M.} \snm{Mora}}
% \author[
%   addressref={aff1},
%   email={fbarranco@ugr.es}
% ]{\inits{FB}\fnm{Francisco} \snm{Barranco}}
% \author[
%   addressref={aff3},
%   email={n.h.abbas@leeds.ac.uk}
% ]{\inits{NA}\fnm{Noorhan} \snm{Abbas}}
% \author[
%   addressref={aff1},
%   email={aguillen@ugr.es}
% ]{\inits{AG}\fnm{Alberto} \snm{Guillén}}
% \author[
%   addressref={aff4},
%   email={o.tsivitanidou@cyens.org.cy}
% ]{\inits{OT}\fnm{Olia} \snm{Tsivitanidou}}
%
% \address[id=aff1]{%
%   \orgdiv{Department of Computer Architecture and Technology},
%   \orgname{University of Granada},
%   \city{Granada},
%   \cny{Spain}
% }
% \address[id=aff2]{%
%   \orgdiv{Department of Signal Theory, Telematics and Communications},
%   \orgname{University of Granada},
%   \city{Granada},
%   \cny{Spain}
% }
% \address[id=aff3]{%
%   \orgdiv{School of Computing},
%   \orgname{University of Leeds},
%   \city{Leeds},
%   \cny{UK}
% }
% \address[id=aff4]{%
%   \orgname{CYENS},
%   \city{Nicosia},
%   \cny{Cyprus}
% }

\end{fmbox}% comment this for two column layout

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                           %%
%% The Abstract begins here                  %%
%%                                           %%
%% Please refer to the Instructions for      %%
%% authors on https://www.biomedcentral.com/ %%
%% and include the section headings          %%
%% accordingly for your article type.        %%
%%                                           %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{abstractbox}

\begin{abstract} % abstract
Messaging platforms are applications, generally mediated by an app, desktop program or the web, mainly used for synchronous communication among users. As such, they have been widely adopted officially by higher education establishments, after little or no study of their impact and perception by the teachers. We think that the introduction of these new tools and the opportunities and challenges they have needs to be studied carefully in order to adopt the model, as well as the tool, that is the most adequate for all parties involved. We already studied the perception of these tools by students, in this paper we examine the teachers’
experiences and perceptions through a survey that we validated with peers, and what they think these tools should make or serve so that it enhances students learning and helps them achieve their learning objectives. The survey has been distributed among tertiary education teachers, both in universitary and other kind of tertiary establishments, based in
Spain (mainly) and Spanish-speaking countries. We have focused on collecting
teachers' preferences and opinions on the introduction of messaging platforms in their day-to-day work, as well as other services attached to them, such as  chatbots. What we intend with this survey is to understand their needs and to gather
information about the various educational use cases where these tools could be
valuable. In addition, an analysis of how and when teachers’ opinions towards
the use of these tools varies across gender, experience, and their discipline
of specialization is presented. The key findings of this study highlight the
factors that can contribute to the advancement of the adoption of messaging
platforms and chatbots in higher education institutions to achieve the desired
learning outcomes.
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% The keywords begin here                  %%
%%                                          %%
%% Put each keyword in separate \kwd{}.     %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{keyword}
\kwd{Chatbots}
\kwd{Messaging platforms}
\kwd{Tutorship}
\kwd{Educational bots}
\kwd{Higher education}
\end{keyword}

\end{abstractbox}
%
%\end{fmbox}% uncomment this for two column layout

\end{frontmatter}

<<setup, echo=FALSE,message=FALSE>>=
knitr::opts_chunk$set(fig.path='figures/')
library(ggplot2)
library(dplyr)
library(tidyr)
library(scales)
library(ggthemes)
library(tidyverse)
library(corrplot)

survey.pilot <- read.csv("data/survey-pilot-3-EN.csv")
survey.pilot %>% group_by( Gender ) %>% summarise ( n = n() ) %>% mutate( freqGender = n/sum(n)) -> freq.Gender
survey.pilot %>% group_by( Age ) %>% summarise ( n = n() ) %>% mutate( freqAge = n/sum(n)) -> freq.Age
survey.pilot %>% group_by( Experience ) %>% summarise ( n = n() ) %>% mutate( freqExperience = n/sum(n)) -> freq.Experience
survey.pilot %>% group_by( Sector ) %>% summarise ( n = n() ) %>% mutate( freqSector = n/sum(n)) -> freq.Sector
survey.pilot %>% group_by( Discipline ) %>% summarise ( n = n() ) %>% mutate( freqDiscipline = n/sum(n)) -> freq.Discipline

survey.pilot %>%  mutate(Messaging.Organization =str_split(Messaging.Organization, ", ")) %>% unnest(cols = c(Messaging.Organization)) -> survey.pilot.MessagingOrganization
survey.pilot.MessagingOrganization[as.character(survey.pilot.MessagingOrganization$Messaging.Organization)!="No", ] -> survey.pilot.MessagingOrganization

survey.pilot %>% group_by( PostCOVID.changes ) %>% filter( n() > 1 ) -> survey.pilot.postCOVID
survey.pilot.postCOVID[as.character(survey.pilot.postCOVID$PostCOVID.changes)!="No", ] -> survey.pilot.postCOVID

survey.pilot %>%  mutate(ChatbotType = ifelse(AgendaBot == "No","","AgendaBot,")) %>%
  mutate(ChatbotType = ifelse(ClassMaterialBot == "No", paste(ChatbotType, "",sep=""), paste(ChatbotType, "ClassMaterialBot,",sep=""))) %>%
  mutate(ChatbotType = ifelse(FAQBot == "No",paste(ChatbotType, "",sep=""), paste(ChatbotType, "FAQBot,",sep=""))) %>%
  mutate(ChatbotType = ifelse(GradesBot == "No", paste(ChatbotType, "",sep=""), paste(ChatbotType, "GradesBot,",sep=""))) %>%
  mutate(ChatbotType = ifelse(OtherBot == "No", paste(ChatbotType, "",sep=""), paste(ChatbotType, "OtherBot,",sep=""))) %>%
  mutate(ChatbotType = substr(ChatbotType,1,nchar(ChatbotType)-1))-> survey.pilot.chatbottype
survey.pilot.chatbottype %>%  mutate(ChatbotType =str_split(ChatbotType, ",")) %>% unnest(cols = c(ChatbotType)) -> survey.pilot.ChatbotType
survey.pilot.ChatbotType[as.character(survey.pilot.ChatbotType$Gender)!="No", ] -> survey.pilot.ChatbotType
survey.pilot.ChatbotType[as.character(survey.pilot.ChatbotType$Experience)!="No", ] -> survey.pilot.ChatbotType

survey.pilot %>%mutate(Discipline = ifelse(as.character(Discipline) == "No", "Other", as.character(Discipline))) %>%
  mutate(Discipline = ifelse(as.character(Discipline) == "Engineering", "Tech/Engineering", as.character(Discipline))) %>%
  mutate(Discipline = ifelse(as.character(Discipline) == "Technology", "Tech/Engineering", as.character(Discipline)))->survey.pilot.Discipline

survey.interaction <- read.csv("data/survey-student-interaction-pilot-3-2021-EN.csv")
survey.interaction[as.character(survey.interaction$Discipline)!="No", ] -> survey.interaction
survey.interaction %>% group_by( Gender ) %>% summarise ( n = n() ) %>% mutate( freqGender = n/sum(n)) -> freqInteraction.Gender
survey.interaction %>% group_by( Age ) %>% summarise ( n = n() ) %>% mutate( freqAge = n/sum(n)) -> freqInteraction.Age
survey.interaction %>% group_by( Experience ) %>% summarise ( n = n() ) %>% mutate( freqExperience = n/sum(n)) -> freqInteraction.Experience

#survey.interaction %>% group_by( Vocational ) %>% summarise ( n = n() ) %>% mutate( freqSector = n/sum(n)) -> freqInteraction.Sector
#freqInteraction.Sector$Sector <- freqInteraction.Sector$Vocational
#freqInteraction.Sector %>% mutate(Sector = ifelse(as.character(Vocational) == "Yes", "Vocational", "University")) -> freqInteraction.Sector
#freqInteraction.Sector <- freqInteraction.Sector[-c(1)]
survey.interaction %>%  mutate(Sector = ifelse(DegreesVocational == "No","","Vocational,")) %>%
  mutate(Sector = ifelse(Sector == "No", paste(Sector, "",sep=""), paste(Sector, "University,",sep=""))) %>%
  mutate(Sector = substr(Sector,1,nchar(Sector)-1))-> df1
df1 %>%  mutate(Sector =str_split(Sector, ",")) %>% unnest(cols = c(Sector)) -> df1.Sector
df1.Sector %>% group_by( Sector ) %>% summarise ( n = n() ) %>% mutate( freqSector = n/sum(n)) -> freqInteraction.Sector

survey.interaction %>% group_by( Discipline ) %>% summarise ( n = n() ) %>% mutate( freqDiscipline = n/sum(n)) -> freqInteraction.Discipline
freqInteraction.Discipline %>%mutate(Discipline = ifelse(as.character(Discipline) == "No", "Other", as.character(Discipline))) -> freqInteraction.Discipline
freq.Discipline %>% mutate(Discipline = ifelse(as.character(Discipline) == "No", "Other", as.character(Discipline))) %>%
                mutate(n = ifelse(as.character(Discipline) =="Engineering", sum(n[as.character(Discipline)=="Engineering" | as.character(Discipline)=="Technology"]), n)) %>%
                mutate(Discipline = ifelse(as.character(Discipline) == "Engineering", "Tech/Engineering", as.character(Discipline))) %>%
                subset(as.character(Discipline) != "Technology")-> freq.Discipline
@

% ********************** INTRODUCTION ****************************

\section{Introduction}

New technologies are being introduced and widely adopted into the classroom in the last years. However, in order to be successful, their application
will require (many) extra work and training for the teacher, as well as for the students. Thus, both of them will acquire new skills and, potentially, pupils will
get a higher engagement to the subject, which is usually the main objective. Even if several technologies are already in use inside the classroom, there is still room for adopting others.
Anyway, this adoption strongly depends on the compromise and capabilities of the teachers, as main active actors in this new academic approach. Given this,
asking and analyzing user's opinions with regard to this scenario will help reaping the expected benefits of the technology.

In the last years, messaging applications such as WhatsApp or Telegram are being used in teaching scenarios mainly at secondary and higher education.
In addition to their main use, some of these tools offer the possibility of interact with \textit{chatbots} (or just bots), even supporting synchronous conversation between the different
parties \cite{differ.chat}. There is also possible to create a standalone chatbot application and attach it to a messaging tool through an API (Application Program Interface).

\textit{Chatbots} are autonomous agents able to communicate and interact with humans (or other chatbots) in a human-like way \cite{Gong2008}, answering question and even posing their own ones to the user.
Thus, we can say that they are close to passing a Turing Test \cite{moor2003turing}. They are implemented by means of different Artificial Intelligence techniques, such as natural language processing.
Inside the plethora of possible applications of chatbots, education is one of the most prolific nowadays \cite{clarizia2018,Smutny2020}. In this domain, they facilitate the so-called
\textit{personalized learning}, i.e. they adapt themselves to students’ pace of learning and provide online tutoring outside the classroom, which have been proved to be very effective increasing their
engagement and participation in the subjects \cite{agarwal2020}. Moreover chatbots' tutoring possibilities aid to reduce the stress of students \cite{Daniel2016,agarwal2020} by avoiding the face-to-face tutoring with their teacher,
or asking questions in front of the class. This situation normally conducts to a silence of the student and a later contact with the teacher by e-mail, which, in turn, will mean a workload for the teacher, that could be very
significant given the existing disproportion between number of students per each teacher.

Hence, the chatbot technology has a potential to mitigate this problem by providing answers to students’ questions and facilitating a dynamic and autonomous learning
experience \cite{Griol2014,kim2020}. In addition, using an automated system such as a chatbot will allow teachers to spend more time on  other topics that students struggle with.

Given the potential benefits that chatbots could bring to the classroom, the
focus of the % EDUBOTS KA2 European project \footnote{https://www.edubots.eu},
unnamed European project, the main sponsor of this study, is to explore best practices and innovative use of chatbots,
and to create a learning community of educators in higher education institutions.
Researchers and teachers participating in this project have developed several studies. Most of these works have been focused on how the
messaging applications and/or chatbots are normally applied to deliver personalized learning in classrooms that occurs anytime anywhere,
promote collaborative learning experiences, group discussions \cite{panah2020study} and boost students' sense of belonging to their institutions
\cite{abbas2021onlinechat}.
The majority of these studies were conducted focusing on the teachers' perspectives, so this paper aims to provide more conclusions regarding
their opinions and needs, as well as the existing challenges or opportunities that arise when adopting these technologies in classrooms. This will lead to clarify
how they can positively contribute to enhancing the learning process in higher education institutions.

In this study, data collection involved two phases:
 \begin{itemize}
\item The first phase aimed to collect feedback from students about their
  students’ preferences of tools/applications for chatting and
  messaging, how they use them in educational context, who they like to be
  with in the class messaging groups, and their expectations of chatbots
  in assisting them during their learning process. Therefore, two surveys
  for bachelor and master degree students at the University of Granada (Spain)
  were designed and answers from more than 250 students were collected. The key
  findings of the students' surveys reveal that they prefer to use messaging
  familiar applications, such as Telegram or WhatsApp.
  Students also expressed interest in chatbots that can assist them with
  organizing their course schedules, help them access their assignment grades
  and facilitate searching for resources.
\item The results of the first phase were fed into the second phase of the
  study, which focused on teachers’ opinions about using chatbots and messaging
  applications in classes. These results are the main focus of this paper.
\end{itemize}

Initially, a single survey was developed for this purpose and responses from 300
higher education teachers were collected and analysed\textcolor{black}{; this will be detailed in Section \ref{sec:meth}}. The responses' analysis
led to the preparation of a second survey for teachers and 200 responses were
collected\textcolor{black}{, again following the methodology that will be explained later. The questions of the second survey focused on asking about teachers'
needs, opinions and preferences of the development of future technology-enhanced
tools and its potential impact on educational institutions policies. In order to assess these, this paper layed out the following research questions}:

\begin{itemize}
\item RQ1 - Do teachers use now or want to use messaging apps in their classes?
\item RQ2 - Which chatbots' features would teachers find useful for this purpose?
\item RQ3 - Which kind of interaction do teachers prefer with their students?
\item RQ4 - Which kind of interaction media features do teachers value the most?
\end{itemize}

We will try to answer these research questions through the analysis of the two aforementioned surveys.
Eventually, what we want to obtain is a series of
recommendations to make a successful deployment of chatbot technologies (and, in
some cases, general instant messaging applications) in higher education; from the response to RQ1 and RQ4 we will try to recommend specific technologies or applications to be used in the classroom; the response to RQ2 and RQ3 will help us recommend chatbots features or specific platforms; and finally, from RQ3 we will also try to find best practices in the adoption of messaging platforms and they accompanying chatbots.

The remainder of the paper is organized as follows: first, an overview of what
current research has found about the use of messaging applications, including
chatbots, in the classroom is described. The methodology used in the surveys is presented in
Section \ref{sec:meth}, and the results of the surveys are presented next in
Section \ref{sec:res}. Finally, we discuss these results and conclude with a
series of recommendations for the successful deployment of chatbots in higher education.

% ********************** STATE OF THE ART ****************************

\section{State of the art}
\label{sec:soa}

% The Use of Mobile Instant Messaging (MIM) Platforms in Higher Education
The widespread and rapid adoption of free Mobile Instant Messaging
(MIM) tools/platforms such as WhatsApp, Telegram, WeChat and Facebook
Messenger stems from their simplicity, ease of use and multi-modality
(i.e. video, audio, text) \cite{tang2017mobile}. Using these tools in
higher education facilitates the delivery of personalized learning
that occurs anytime anywhere, and promotes collaborative learning
experiences and group discussions \cite{panah2020study}.

WhatsApp is, at least in most Western countries, the most popular MIM platform
used by educators to provide assignments' feedback to students, support course
discussions, and contribute resources in informal learning settings
\cite{panah2020study}.  Moreover, the use of WhatsApp in higher education could
enhance social presence \cite{tang2017mobile} and foster trust relationships
between educators and students embedded in the social learning process
\cite{gachago2015crossing}. However, this last paper also reflects the need for
learners to "take ownership of the tool" and the advantages of social learning
in general. At the same time, it also mentions different challenges, among which
the most important is the blurring of social and academic life; indeed, there are challenges when using MIM tools that occur due to the
blurring of boundaries between academic and private life. This can lead to
technostress \cite{gachago2015crossing}, difficulty in managing
responsibilities, especially among mature students, and lack of privacy
\cite{tang2017mobile}.  Students' dropout of the MIM groups, as they can leave
groups at any time, can hinder their learning and undermine educators' efforts
\cite{mwakapina2016whatsapp}.  In addition, there is a need to set rules and
norms for these MIM groups in order to maintain the safety of these online
communities for students \cite{abbas2021onlinechat}. However, these rules should
not affect students' ownership and control, since it is vital to advance in
their learning \cite{gachago2015crossing} process.  This is why examining the
role of MIM platforms in higher education is still a challenge, and why the opinions of
the teaching community towards them have to be examined, as we do in this paper.

% The Use of Chatbots in Higher Education
The use of MIM platforms, although possibly valuable by itself, can be enhanced with the
use of chatbots, which are conversational agents that normally dwell 
synchronous conversations systems. The use of chatbots in higher education is still in its early stages
\cite{yang2019opportunities}. Nevertheless, recent studies have shown their
positive impact on students' academic performance \cite{perez2020rediscovering}
and engagement \cite{differ.chat,abbas2021onlinechat}, leading to a growing
interest in using this technology in the (possibly virtual) classroom. Indeed,
the using chatbots to collect course feedback from students in higher
education improved response quality and boosted enjoyment levels
\cite{abbas2021Surveys}.  According to \cite{Roblyer2010}, the use of
mobile devices along with gamification strategies
\cite{Yildirim2017} can improve student motivation. In this sense, the authors
in \cite{Pimmer2019} adopted a quasi-experimental, survey-based approach to
report the positive impact of using instant messaging tools in boosting
students' knowledge and mitigating their feelings of isolation.

Several studies have been conducted to evaluate chatbots in higher education. For
instance, a recent study was presented by Smutny and Schreiberova
\cite{Smutny2020} reviewing 47 educational chatbots implemented in Facebook
Messenger, with a focus on identifying characteristics and quality metrics such as
language, subject matter, and platform. Pérez and collaborators
\cite{perez2020rediscovering} attempted to categorize educational chatbots attending to their purpose 
as service-oriented or teaching-oriented. Service-oriented chatbots
include those that provide service support, such as Ask Holly
\cite{meetholly} and Dina \cite{santoso2018dinus}, which responds to
students' questions about enrollment and registration. Ask L. U.
\cite{LancasterUniversity_2019} answers students' frequently asked questions
about schedules, grades, tutors, and societies. LISA \cite{dibitonto2018chatbot}
and Differ \cite{differ.chat} facilitate breaking the ice between new students
by introducing them to each other. Ranoliya et al. \cite{ranoliya2017chatbot}
proposed a generic chatbot for university students that is able to answer 
frequently asked questions. The University of Granada had
Elvira, a built-in chatbot in its main web page \cite{MOREO20129} to perform the same task. In addition to being able
to answer pre-established frequent questions, it did so from the website of
the University of Granada using an inset persona, voiced by a real person, who
lip-synced the answers. However, in the case of Elvira, the emphasis was on the
authenticity of the synchronization rather than making the updated 
of answered questions simpler, more interactive, or more open to the rest of the university staff using actual messaging platforms. In
most cases, the embedded search engine was able to provide more up-to-date
and accurate answers than those provided by Elvira, which had to be updated
by hand (and not too often). It was eventually discontinued, and its technology was not adapted to other platforms, since it did not really provide a useful service. Its service was not tied to teaching anyway; it was more related to directory queries and university-wide administrative questions, so even if it is strictly an example of a chatbot in a university setting, it is not really a chatbot that can be adopted by teachers to affect the learning outcomes, which is the focus of this paper.

On the other hand, teaching-oriented chatbots are more sophisticated, as they
set personalized learning outcomes and monitor learning progress. For instance,
\cite{fernoagua2018intelligent} reported on ``eduAssistant'', a virtual teaching
assistant chatbot developed on the Telegram messaging platform. In this study,
the Telegram platform was chosen because it is easy to use, students are familiar
with its features, and it enables them to exchange messages in different formats
(text, audio and video) \cite{fernoagua2018intelligent}. In addition, Telegram
can operate on all devices and operating systems. The ``eduAssistant'' chatbot
acts as an automatic agent in the teacher-content-student relationship, providing real-time
feedback loops and a personalized learning experience relevant to the
students' acquired skills and knowledge. Using this chatbot, educators can
create interactive instances in their lectures where they pose questions to
their students and the chatbot assists those who need further help by giving
them more hints and reporting it to the educator's dashboard
\cite{fernoagua2018intelligent}. This can help educators identify
students who need more attention and provide them with more educational resources
relevant to their academic achievement. %In addition, a recent study
%has analysed how the use of chatbots positively affected the learning outcomes
%of students in a Chinese class \cite{doi:10.1177/0735633120929622}.

Despite its (arguably) successful implementation in different higher education
institutions, particularly on the fringes of educational activity rather than actual
student-teacher interaction, its implementation or deployment is not trivial.
Some authors, for instance \cite{sjostrom2018designing}, have proposed a
conceptual architecture for the adoption of teaching-oriented chatbots in higher
education.  This conceptual architecture is based on a systematic literature
review of previous studies examining the design of chatbots in higher
education, as well as making a content analysis of student emails and discussion
forum posts from four instances of a Java programming course.  The study outlined
several design considerations; among them, the authors emphasized the importance
of developing chatbots in platforms that students and educators are familiar
with and can easily access (i.e. Facebook Messenger), which was confirmed by
\cite{hobert2019you} and \cite{fernoagua2018intelligent}. In addition, \cite{sjostrom2018designing} argued that 
conceptualizing learners' questions could help designers integrate the
appropriate types of questions that chatbots should support for different
courses. Other authors, such as \cite{coronado2018cognitive}, have proposed agents that
store learning materials to be provided on demand to students, while \cite{crockett2017predicting} 
reported on tutoring systems that can
perform initial assessments of students' understanding and provide learning
materials to advance their understanding to the next level.

Regarding the factors for the adoption of chatbots in higher education, many
studies have focused on the evaluation of technology acceptance and usability
\cite{Roblyer2010,Pimmer2019}. However, higher education is a special domain
where, according to \cite{hobert2019you}, specific pedagogical factors
such as learning success and increased motivation are more important.  Therefore,
to develop effective chatbots for higher education, the needs of all stakeholders
(i.e. educators, students, institutions, etc) should be carefully
collected and taken into consideration
\cite{sjostrom2018designing,tsivitanidouusers}. These needs include, but are not limited to, student learning success due to higher motivation, but these are a posteriori effects that cannot be assessed in advance. This is why we focus in this paper on what  teachers are looking for in a Technology Acceptance Model (TAM) and associated technologies like chatbots. Both authors focus on what teachers need in terms of the latter; this paper will focus on a wider perspective, trying to determine what they are looking for in terms of general messaging technology. The classical literature \cite{tam91} already proposes that any adoption of technology must be tuned to the user needs and experience. In this paper, we will try to find out what those are in order to propose a successful model of adoption of chatbot technology.

In line with this research, our previous work \cite{MoraChatbots2021:anon}
aimed at analyzing the expectations of students in this regard. Recent papers
have also analysed how the use of chatbots affected the learning outcomes of
students in a Chinese class \cite{doi:10.1177/0735633120929622}. This paper
analysed conversational chatbots in a one-to-one setting, finding that learning
truly benefited from it. This is, once again, an evidence of the benefit of
chatbots in certain settings; however, there are some prior experiences, as well as needs, that may
prevent the successful acceptance of the technology, and these are what we are
trying to find out in this paper, alng with what kind of features would improve
its acceptance.

Besides, this work is focused on the other key actor in this challenge: 
educators. How they accept chatbots has been studied very recently by Chocarro et
al. \cite{doi:10.1080/03055698.2020.1850426}, who, by analyzing surveys, created
a TAM that proposed a series of features that
would make chatbots easier to accept, including formality of language as well
as easiness and usefulness. The survey targeted primary and secondary education
teachers and was also more interested in the general use of chatbots in
education, not specifically in a classroom setting, seeking educational
outcomes. However, their results are obviously
interesting and relevant for this work.

We will present next the methodology we followed to find out teachers' opinions.

% ********************** METHODOLOGY ****************************

\section{Methodology}
\label{sec:meth}

\textcolor{black}{This study followed a quantitative approach towards addressing its research objective, by collecting rich data about the use of messaging applications and chatbots
by educators in both universities and colleges\footnote{In Spain, they are called simply high schools, and deliver
tertiary or associated degrees that do not have the consideration of university degrees}.
To achieve this goal two online surveys were designed and developed using Google Forms
in Spanish. They are comprised of six questions focusing on demographic data
(i.e., sector, gender, degrees, discipline, age, teaching experience)
followed by several multiple choice questions allowing the participants to choose multiple answers.}

In the {\bf first survey}, the multiple choice questions are
focused on the use of messaging apps in the teaching practice, the type of chatbot use cases that teachers consider
useful for their teaching and the impact of COVID on the teaching practice; please check them out to the full extent in the Appendix.\footnote{This {\bf first survey} was made in two separate Google forms, one for
university and the other for teachers in tertiary education institutions, outside the university system (vocational training, including 1- and 2-year degrees). Questions and
responses were the same, except for the type of tertiary degrees that
were considered.}

\textcolor{black}{This survey was piloted by the authors of the paper and their colleagues before using it in the
study.  All the comments collected from this trial were incorporated in the survey.
We mainly used mailing lists, as well as Telegram groups,
to post the survey link to reach out to more educators. The university form was sent to Spanish
(mainly Andaluc\'{\i}a and Galicia) universities, and also  universities in Costa Rica and
Mexico. The tertiary (non-university/college) teachers were based mainly in
Andaluc\'{\i}a. Dissemination and answers took place in the first quarter of
2021, during the Covid-19 pandemic and while, at least in Spain, many universities had
mandatory virtual teaching.}

A total of 282 teachers responded to the survey: 193 teaching at the University
(68.4\%) and 89 at other tertiary education institutions (31.6\%). \textcolor{black}{From those, 179 indicated their gender as male (63.5\%), 98 as female (34.8\%) and 5 teachers preferred not to indicate their gender (1.8\%). In terms of age, the majority of the participants (n=111) in this survey was 45-55 years old (39.4\%), while 91 teachers were 35-45 (32.3\%), 46 were 25-35 (16.3\%) and 34 teachers were older than 55 (12.1\%). Last, 84 teachers had a teaching experience of 16-25 years (29.8\%), 75 teachers 6-15 years of experience (26.6\%), 69 teachers had 0-5 years of experience (24.5\%) and finally, 54 teachers had more than 25 years of teaching experience (19.1\%).}

Responses were stored automatically in a Google Drive spreadsheet. Eventually,
the results from the two forms we used for the first survey were collated in a
single spreadsheet. Survey questions are shown in Appendix I (Section \ref{app:questions}).

The {\bf second survey} was designed after initial results to the first survey arrived,
and pointed to necessities and experiences of teachers not being addressed by the first one; namely
their experience with messaging platforms and the way they were used to interact with students.
It was piloted in a group of university teachers attending to a formation course, and validated with them.
It was then extended to the rest of the responders, using the same media: Telegram groups, email and announcement in mailing lists. The questions asked in that survey are also shown in Appendix I
(Section \ref{app:questions}).

A total of 205 teachers responded to the second survey: 187 graduate teachers (91.2\%) and 18 student teachers (8.8\%). From those, 124 were male (60.5\%), 65 female (31.7\%) and 16 teachers preferred not to indicate their gender (7.8\%). In terms of age, the majority of the participants (n=70) was again, as in the first survey, 45-55 years old (34.1\%), while 67 teachers were 35-45 years old (32.7\%), 42 were 25-35 (20.5\%) and 26 teachers were older than 55 (12.7\%). In terms of teaching experience, 59 teachers had a teaching experience of 16-25 years (28.8\%), 51 teachers had 6-15 years (24.9\%) and 0-5 years of experience (24.9\%), and 44 teachers more than 25 years of teaching experience (21.5\%).

\subsection{Survey validation}

To avoid any bias and to be sure about the quality of the questionnaire, a panel board was consulted. There were three questions for each of the items of the questionnaire which were: Clarity, Importance, Adequateness.
Clarity was measured as an answer to the question: Is the item well presented and without ambiguity? \footnote{Original text: ¿El ítem está bien redactado y sin ambigüedad?} with two choices: Yes or No.
Both Importance and Adequateness were measured in a Likert scale from 1 to 4 (1 low, 4 high).

The panel board was composed by 7 volunteer experts that have wide experience as educators and researchers. The results obtained are summarized in Table \ref{tab:validation}.

As validation shows, all expert agree without any doubt about the clarity of each question. Furthermore, there is a large consensus regarding the Importance and Adequateness of all the eight questions in the questionnaire.

\begin{table}
\begin{tabular}{|c|ccc|}
    \hline
    Questions & Clarity (Yes/No) & Avg. Importance & Avg. Adequateness \\
    \hline
    Q1 &  Yes (100\%)  &  3,571 (0,53) & 3,571 (0,53) \\
    Q2 &  Yes (100\%)  &  3,714 (0,49) & 3,571 (0,53) \\
    Q3 &  Yes (100\%)  &  3,857 (0,38) & 3,857 (0,38) \\
    Q4 &  Yes (100\%)  &  3,571 (0,79) & 3,857 (0,38) \\
    Q5 &  Yes (100\%)  &  3,714 (0,49) & 3,714 (0,49) \\
    Q6 &  Yes (100\%)  &  3,714 (0,49) & 3,857 (0,38) \\
    Q7 &  Yes (100\%)  &  3,714 (0,49) & 3,714 (0,49) \\
    Q8 &  Yes (100\%)  &  4 (0) & 4,000 (0) \\
    \hline
\end{tabular}
\caption{Validation results from the panel board (7 experts) for each question in the research questionnaire. The range of values for Importance and Adequateness is from 1 to 4 (higher is better).}\label{tab:validation}
\end{table}

Next, we show the results and analyse them.

% ********************** RESULTS ****************************

\section{Results and analysis}
\label{sec:res}

% Describe the survey and population
We collected responses from 282 teachers from Spain and Spanish Speaking
countries for \textit{survey1} and 205 for \textit{survey2}. The two forms were
open for approximately the same amount of time, around two months. The sample
includes mostly teachers from university, although about 32\% and 8\% are from
non-universitary tertiary education in the first and second survey respectively. Regarding
gender, 61\% of the teachers were male and 32\% female, while approximately 5\%
chose not to disclose it. Finally, the responses are more or less equally
distributed according to the teaching experience, showing 24\% of responses from
teachers with 5 or less years of experience, 26\% for 6-15 years of experience,
30\% for 16-25, and 20\% for teachers with more than 25 years of experience.

We will try to find the answers to the four research questions next, by
analyzing the response to the different survey questions.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% RQ1 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{RQ1 - Are teachers already using messaging apps in their classes?}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Group messaging apps have become a powerful tool for \textcolor{black}{communicating} with
students. Also, many of them have built-in chatbots to enhance the learning
process. After an initial analysis to the answers to the first survey, we
realized that there were some prior issues mainly related to the adoption of a
technology, \textcolor{black}{i.e. chatbots}, that generally piggybacks on another, \textcolor{black}{i.e. messaging
applications. Generally,} in a technology adoption model, perceived ease of use and
perceived usefulness are essential. But it is going to be very difficult for
teachers to find {\em chatbots} easy to use if they, previously, do not
consider, or simply use, the messaging tools to which they are attached.

This is why for answering RQ1, teachers were queried about whether they use
messaging apps in their classes, specifically, Telegram, WhatsApp, Slack (an
application used mainly in IT departments and software development), or any
other messaging app, and whether they use messaging apps provided by their
academic institution (see Table \ref{tab:use_messapps}).
Overall, the majority of the teachers responded that they do use messaging apps
in their classrooms, from which apps (or, more probably, messaging solutions) provided by the academic institution
(n=159, 56\%) and WhatsApp (n=124, 44.0\%) were the most common responses. Only
19 teachers (6.7\%) replied that they do not use any messaging app in their
class.
%
\begin{table}[h!tbp]
\begin{tabular}{lllll}
\hline
\multicolumn{1}{c}{Messaging App} & \multicolumn{2}{c}{Yes} & \multicolumn{2}{l}{No} \\ \cline{2-5}
\multicolumn{1}{c}{} & \multicolumn{1}{c}{Frequency} & \multicolumn{1}{c}{\%} & Frequency & \multicolumn{1}{c}{\%} \\ \hline
Telegram & 62 & 22.0 & 220 & 78.0 \\
WhatsApp & 124 & 44.0 & 158 & 56.0 \\
Slack & 15 & 5.3 & 267 & 94.7 \\
Other & 60 & 21.3 & 222 & 78.7 \\
\begin{tabular}[c]{@{}l@{}}Provided by the Academic\\ Institution\end{tabular} & 159 & 56.4 & 123 & 43.6 \\
None & 19 & 6.7 & 263 & 93.3 \\ \hline
\end{tabular}
\vspace{0.8em}
\caption{Use of messaging apps to assist the learning process.}
\label{tab:use_messapps}
\end{table}

% Fran: I think figures with data about the use of messaging apps should come here. Then, we say we also asked about changes after covid and include data about that
Results in Figs. \ref{figure:provided*chatbots_use},
\ref{figure:whatsapp*chatbots_use}, \ref{figure:telegram*chatbots_use} do not
show significant differences in the use of instant messaging apps between
teachers from University and vocational education. In general, most teachers
prefer messaging apps provided by their own institutions (we also show the
results for the two most popular messaging app platforms: WhatsApp and
Telegram). With respect to specific disciplines, Engineering and Technology
teachers are more active in their use, but the number of teachers from Humanities
who answered they used these apps in their classes is also remarkable
(around 60\% use the apps provided by their institutions).

Although no significant differences were found regarding gender, female teachers
answered they use instant messaging apps more than male teachers (about 10\%
more). Also, teachers in vocational Education use WhatsApp more than university
teachers. Regarding the distribution of the use of messaging apps per age, there
are no significant differences for WhatsApp and apps provided by their own
academic institutions. % Olia requests support for this.
However, younger teachers also use Telegram with more than 25\% responding they do, a percentage
that falls to about 10\% for teachers that are 55 or older. %The third column
                                %compares the use of these apps considering the
                                %number of years of experience.
One interesting result is that about 65\% of teachers with more than 25 years of
experience use the platforms provided by their institutions while the percentage
goes down to less than 50\% for teachers with 6-15 years of experience.
%
\begin{figure}[h!tbp]
\begin{center}
    \begin{minipage}[t]{0.35\textwidth}
    \centering
<<gender_provided, echo=FALSE, fig.width=10/2.54, fig.height=7/2.54>>=
survey.pilot[as.character(survey.pilot$Gender)!="No", ] -> df1
df1[as.character(df1$Gender)!="PNTS", ] -> df1
ggplot(df1, aes(x = MessagingPlatformProvided, group = Gender)) + geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count", show.legend = FALSE) + scale_y_continuous(labels=scales::percent) + ylab("relative frequencies") + facet_grid(~factor(Gender, levels=c('Female','Male')))+ theme(axis.title.y=element_blank()) + theme(axis.title.x=element_blank())
#+ facet_grid(~factor(Gender, levels=c('Female','Male','PNTS')))+ theme(axis.title.y=element_blank())
@
%       \includegraphics[width=\textwidth]{figures/gender_provided.pdf}
    \end{minipage}
    \begin{minipage}[t]{0.35\textwidth}
    \centering
<<experience_provided, echo=FALSE, fig.width=10/2.54, fig.height=7/2.54>>=
survey.pilot[as.character(survey.pilot$Experience)!="No", ] -> df1
ggplot(df1, aes(x = MessagingPlatformProvided, group = Experience)) + geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count", show.legend = FALSE) + scale_y_continuous(labels=scales::percent) + ylab("relative frequencies") + facet_grid(~factor(Experience, levels=c("0-05 years","06-15 years", "16-25 years", "> 25 years")))+ theme(axis.title.y=element_blank()) + theme(axis.title.x=element_blank())
@
%       \includegraphics[width=\textwidth]{figures/experience_provided.pdf}
    \end{minipage}
    \begin{minipage}[t]{0.3\textwidth}
        \centering
<<sector_provided, echo=FALSE, fig.width=10/2.54, fig.height=8.25/2.54>>=
ggplot(survey.pilot, aes(x = MessagingPlatformProvided, group = Sector)) + geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count", show.legend = FALSE) + scale_y_continuous(labels=scales::percent) + ylab("relative frequencies") + facet_grid(~Sector)+ theme(axis.title.y=element_blank())+ theme(axis.title.x=element_blank())
@
%       \includegraphics[width=\textwidth]{figures/sector_provided.pdf}
    \end{minipage}
    %
    \begin{minipage}[t]{0.6\textwidth}
        \centering
<<discipline_provided, echo=FALSE, fig.width=18.6/2.54, fig.height=7.7/2.54>>=
survey.pilot.Discipline[as.character(survey.pilot.Discipline$Discipline)!="Other", ] -> survey.pilot.Discipline
ggplot(survey.pilot.Discipline, aes(x = MessagingPlatformProvided, group = Discipline)) + geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count", show.legend = FALSE) + scale_y_continuous(labels=scales::percent) + ylab("relative frequencies") + #facet_grid(~factor(Discipline, levels = c("Biomedicine", "Humanities", "Social Sciences", "Sciences", "Tech/Engineering", "Other")))+ theme(axis.title.y=element_blank())
facet_grid(~factor(Discipline, levels = c("Biomedicine", "Humanities", "Social Sciences", "Sciences", "Tech/Engineering")))+ theme(axis.title.y=element_blank())+ theme(axis.title.x=element_blank())
@
%       \includegraphics[width=\textwidth]{figures/discipline_provided.pdf}
    \end{minipage}
\end{center}
    \caption{Use of messaging platforms either external or provided by the teachers' academic institutions in class: distributions per gender, years of experience in education, university or vocational education, and discipline.}
    \label{figure:provided*chatbots_use}
\end{figure}
%
\begin{figure}[h!tbp]
\begin{center}
    \begin{minipage}[t]{0.35\textwidth}
    \centering
<<gender_whatsapp, echo=FALSE, fig.width=10/2.54, fig.height=7/2.54>>=
survey.pilot[as.character(survey.pilot$Gender)!="No", ] -> df1
df1[as.character(df1$Gender)!="PNTS", ] -> df1
ggplot(df1, aes(x = MessagingPlatformWhatsApp, group = Gender)) + geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count", show.legend = FALSE) + scale_y_continuous(labels=scales::percent) + ylab("relative frequencies") + facet_grid(~factor(Gender, levels=c('Female','Male')))+ theme(axis.title.y=element_blank()) + theme(axis.title.x=element_blank())
@
%       \includegraphics[width=\textwidth]{figures/gender_whatsapp.pdf}
    \end{minipage}
    \begin{minipage}[t]{0.35\textwidth}
    \centering
<<experience_whatsapp, echo=FALSE, fig.width=10/2.54, fig.height=7/2.54>>=
survey.pilot[as.character(survey.pilot$Experience)!="No", ] -> df1
ggplot(df1, aes(x = MessagingPlatformWhatsApp, group = Experience)) + geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count", show.legend = FALSE) + scale_y_continuous(labels=scales::percent) + ylab("relative frequencies") + facet_grid(~factor(Experience, levels=c("0-05 years","06-15 years", "16-25 years", "> 25 years")))+ theme(axis.title.y=element_blank()) + theme(axis.title.x=element_blank())
@
%       \includegraphics[width=\textwidth]{figures/experience_whatsapp.pdf}
    \end{minipage}
    \begin{minipage}[t]{0.3\textwidth}
    \centering
<<sector_whatsapp, echo=FALSE, fig.width=10/2.54, fig.height=8.25/2.54>>=
ggplot(survey.pilot, aes(x = MessagingPlatformWhatsApp, group = Sector)) + geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count", show.legend = FALSE) + scale_y_continuous(labels=scales::percent) + ylab("relative frequencies") + facet_grid(~Sector)+ theme(axis.title.y=element_blank()) + theme(axis.title.x=element_blank())
@
%       \includegraphics[width=\textwidth]{figures/sector_whatsapp.pdf}
    \end{minipage}
    \begin{minipage}[t]{0.6\textwidth}
    \centering
<<discipline_whatsapp, echo=FALSE, fig.width=18.6/2.54, fig.height=7.7/2.54>>=
ggplot(survey.pilot.Discipline, aes(x = MessagingPlatformWhatsApp, group = Discipline)) + geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count", show.legend = FALSE) + scale_y_continuous(labels=scales::percent) + ylab("relative frequencies") + facet_grid(~factor(Discipline, levels = c("Biomedicine", "Humanities", "Social Sciences", "Sciences", "Tech/Engineering")))+ theme(axis.title.y=element_blank()) + theme(axis.title.x=element_blank())
@
%       \includegraphics[width=\textwidth]{figures/discipline_whatsapp.pdf}
    \end{minipage}
\end{center}
    \caption{Use of Whatsapp in class: distributions per gender, years of experience in education, university or vocational education, and discipline.}
    \label{figure:whatsapp*chatbots_use}
\end{figure}
%
\begin{figure}[h!tbp]
\begin{center}
    \begin{minipage}[t]{0.35\textwidth}
    \centering
<<gender_telegram, echo=FALSE, fig.width=10/2.54, fig.height=7/2.54>>=
survey.pilot[as.character(survey.pilot$Gender)!="No", ] -> df1
df1[as.character(df1$Gender)!="PNTS", ] -> df1
ggplot(df1, aes(x = MessagingPlatformTelegram, group = Gender)) + geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count", show.legend = FALSE) + scale_y_continuous(labels=scales::percent) + ylab("relative frequencies") + facet_grid(~factor(Gender, levels=c('Female','Male')))+ theme(axis.title.y=element_blank()) + theme(axis.title.x=element_blank())
@
%       \includegraphics[width=\textwidth]{figures/gender_telegram.pdf}
    \end{minipage}
    \begin{minipage}[t]{0.35\textwidth}
    \centering
<<experience_telegram, echo=FALSE, fig.width=10/2.54, fig.height=7/2.54>>=
survey.pilot[as.character(survey.pilot$Experience)!="No", ] -> df1
ggplot(df1, aes(x = MessagingPlatformTelegram, group = Experience)) + geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count", show.legend = FALSE) + scale_y_continuous(labels=scales::percent) + ylab("relative frequencies") + facet_grid(~factor(Experience, levels=c("0-05 years","06-15 years", "16-25 years", "> 25 years")))+ theme(axis.title.y=element_blank()) + theme(axis.title.x=element_blank())
@
%       \includegraphics[width=\textwidth]{figures/experience_telegram.pdf}
    \end{minipage}
    \begin{minipage}[t]{0.3\textwidth}
    \centering
<<sector_telegram, echo=FALSE, fig.width=10/2.54, fig.height=8.25/2.54>>=
ggplot(survey.pilot, aes(x = MessagingPlatformTelegram, group = Sector)) + geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count", show.legend = FALSE) + scale_y_continuous(labels=scales::percent) + ylab("relative frequencies") + facet_grid(~Sector)+ theme(axis.title.y=element_blank()) + theme(axis.title.x=element_blank())
@
%       \includegraphics[width=\textwidth]{figures/sector_telegram.pdf}
    \end{minipage}
    \begin{minipage}[t]{0.6\textwidth}
    \centering
<<discipline_telegram, echo=FALSE, fig.width=18.6/2.54, fig.height=7.7/2.54>>=
ggplot(survey.pilot.Discipline, aes(x = MessagingPlatformTelegram, group = Discipline)) + geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count", show.legend = FALSE) + scale_y_continuous(labels=scales::percent) + ylab("relative frequencies") + facet_grid(~factor(Discipline, levels = c("Biomedicine", "Humanities", "Social Sciences", "Sciences", "Tech/Engineering")))+ theme(axis.title.y=element_blank()) + theme(axis.title.x=element_blank())
@
%       \includegraphics[width=\textwidth]{figures/discipline_telegram.pdf}
    \end{minipage}
\end{center}
    \caption{Use of Telegram in class: distributions per gender, years of experience in education, university or vocational education, and discipline.}
    \label{figure:telegram*chatbots_use}
\end{figure}

% This is about changes (due to) after covid
Some of the questions in the survey were focused on the impact of the COVID19
pandemic between the 2020 and 2021 academic years in the teachers' attitudes towards the use of
instant messaging apps in their class. Our main intention in this case was to
assess whether a crisis will bring about some kind of change in the use of
tools. The 282 answers are summarized in
Figs. \ref{figure:gender*chatbots_use_postcovid},
\ref{figure:sector*chatbots_use_postcovid},
\ref{figure:years*chatbots_use_postcovid},
\ref{figure:disc*chatbots_use_postcovid}, showing that about 77\% of teachers
already used these tools before the pandemic and kept using them during the
pandemic lockdowns that forced students and educators to use remote education
schemes. Moreover, approximately 15\% of them switched their messaging app for
one that offered a safer interaction with their students. According to the
responses, an additional 16\% started using messaging apps during the pandemic
for the first time in their classes. % Aggregating the answers, currently 91\% of
% the teachers are using messaging apps as learning tools.
%A few relevant facts from the collected data: teachers with 16-25 years of experience seem more worried about the security of messaging apps and teachers with up to 5 years of experience were more open to adopt messaging apps due to the pandemic; female teachers were more willing to adopt new platforms due to the pandemic, and even to start using safer alternatives; finally, university teachers that did not use them were more reluctant to start using them after the pandemic.
\begin{figure}[h!tbp]
\begin{center}
    \begin{minipage}[t]{0.80\textwidth}
        \centering
<<gender_covid, echo=FALSE, fig.width=19/2.54, fig.height=5.5/2.54>>=
survey.pilot.postCOVID <- within(survey.pilot.postCOVID, PostCOVID.changes <- factor(PostCOVID.changes, levels=names(sort(table(PostCOVID.changes), decreasing=TRUE))))
ggplot(survey.pilot.postCOVID, aes( x = PostCOVID.changes, y = ..count.., group=Gender, fill = Gender ) ) + geom_bar()+ theme(axis.text.x = element_text(angle = 90), size=7) +coord_flip()+ theme_tufte()+ theme(axis.title.y=element_blank())
@
%       \includegraphics[width=\textwidth]{figures/gender_covid.pdf}
        \vspace{-0.5cm}
        \caption{Total count of responses for the use of messaging apps after the COVID-19 pandemic grouped by gender (PNTS stands for Prefer Not To Say).}
        \label{figure:gender*chatbots_use_postcovid}
    \end{minipage}
    \begin{minipage}[t]{0.80\textwidth}
        \centering
<<sector_covid, echo=FALSE, fig.width=19/2.54, fig.height=5.5/2.54>>=
ggplot(survey.pilot.postCOVID, aes( x = PostCOVID.changes, y = ..count.., group=Sector, fill = Sector ) ) + geom_bar()+ theme(axis.text.x = element_text(angle = 90)) +coord_flip()+ theme_tufte()+ theme(axis.title.y=element_blank())
@
%       \includegraphics[width=\textwidth]{figures/sector_covid.pdf}
        \vspace{-0.5cm}
        \caption{Total count of responses for the use of messaging apps after the COVID-19 pandemic grouped by sector.}
        \label{figure:sector*chatbots_use_postcovid}
    \end{minipage}
    \begin{minipage}[t]{0.80\textwidth}
        \centering
<<experience_covid, echo=FALSE, fig.width=19/2.54, fig.height=5.5/2.54>>=
ggplot(survey.pilot.postCOVID, aes( x = PostCOVID.changes, y = ..count.., group=Experience, fill = Experience ) ) + geom_bar()+ theme(axis.text.x = element_text(angle = 90)) +coord_flip()+ theme_tufte()+ theme(axis.title.y=element_blank())
@
%       \includegraphics[width=\textwidth]{figures/experience_covid.pdf}
        \vspace{-0.5cm}
        \caption{Total count of responses for the use of messaging apps after the COVID-19 pandemic grouped by years of experience in teaching.}
        \label{figure:years*chatbots_use_postcovid}
    \end{minipage}
        \begin{minipage}[t]{0.80\textwidth}
        \centering
<<discipline_covid, echo=FALSE, fig.width=19/2.54, fig.height=5.5/2.54>>=
survey.pilot.postCOVID %>%
  mutate(Discipline = ifelse(as.character(Discipline) == "Engineering", "Tech/Engineering", as.character(Discipline))) %>%
  mutate(Discipline = ifelse(as.character(Discipline) == "Technology", "Tech/Engineering", as.character(Discipline)))->survey.pilot.postCOVID
survey.pilot.postCOVID[as.character(survey.pilot.postCOVID$Discipline)!="No", ] -> survey.pilot.postCOVID
ggplot(survey.pilot.postCOVID, aes( x = PostCOVID.changes, y = ..count.., group=Discipline, fill = Discipline ) ) + geom_bar()+ theme(axis.text.x = element_text(angle = 90)) +coord_flip()+ theme_tufte()+ theme(axis.title.y=element_blank())
@
%       \includegraphics[width=\textwidth]{figures/sector_covid.pdf}
        \vspace{-0.5cm}
        \caption{Total count of responses for the use of messaging apps after the COVID-19 pandemic grouped by discipline.}
        \label{figure:disc*chatbots_use_postcovid}
    \end{minipage}
\end{center}
    % Have different sub-captions so that we can refer to specific subfigures - Olia
\end{figure}

A chi-square test of independence was performed to examine the relation between instructors’ discipline, sector, and gender and any potential changes that occurred to the use of messaging apps due to the COVID-19 pandemic. The relation between the latter variable and the instructors’ sector was significant, $X^{2}_{(4, N = 282)} = 9.598$, $p =0.048$. Table \ref{tab:sector*covid} shows the frequencies that are cross-tabulated vs. changes undertaken after the COVID pandemic. This finding suggests that how teachers used messaging apps during the pandemic, specifically whether they changed their habits for teaching purposes, was related to the sector (university vs vocational) in which they were working.
%
\begin{table}[h!tbp]
\begin{tabular}{llccc}
\hline
 & \multicolumn{1}{c}{} & \multicolumn{2}{c}{Sector} & Total \\ \cline{3-5}
 & \multicolumn{1}{c}{} & \begin{tabular}[c]{@{}c@{}}University \\ (f)\end{tabular} & \begin{tabular}[c]{@{}c@{}}Vocational\\ (f)\end{tabular} &  \\ \cline{2-5}
\multirow{4}{*}{\begin{tabular}[c]{@{}l@{}}Post-covid\\ changes\end{tabular}} & \begin{tabular}[c]{@{}l@{}}Yes, I use a safer messaging \\ app now\end{tabular} & 27 & 13 & 40 \\
 & \begin{tabular}[c]{@{}l@{}}Yes, I didn’t use any \\ messaging app before, but \\ do use it now\end{tabular} & 31 & 13 & 44 \\
 & \begin{tabular}[c]{@{}l@{}}No, I did use it before and \\ use it now\end{tabular} & 105 & 60 & 165 \\
 & \begin{tabular}[c]{@{}l@{}}No, I didn’t use it before and \\ don’t use it now\end{tabular} & 22 & 2 & 24 \\
Total &  & 193 & 89 & 282 \\ \hline
\end{tabular}
\vspace{0.8em}
\caption{Sector * post-covid changes cross-tabulation}
\label{tab:sector*covid}
\end{table}

The majority of the teachers (n=165) mentioned that no changes in their habits occurred due to the emergency remote teaching, as the use of messaging apps was part of their teaching practices and remains the same, from which 105 teachers come from the university, and 60 teachers from the non-universitary tertiary sector. The relationships between changes in the use of messaging apps, due to the covid-19 pandemic, and the instructors’ discipline, $X^{2}_{(24, N = 282)} = 44.856$, $p =0.006$, as well as the gender, $X^{2}_{(8, N = 282)} = 16.249$, p= 0.039 were also significant. This finding indicates that how teachers responded to the use of messaging apps during the pandemic was also related to their gender and discipline. In fact, from the majority of the teachers who did not change their habits in this respect (n=165), most of them are males (n=114), and come from the technology (n=60) and engineering (n=53) disciplines (see Tables \ref{tab:gender*covid}, \ref{tab:discipline*covid}).
%
\begin{table}[h!tbp]
\begin{tabular}{llcccc}
\hline
 & \multicolumn{1}{c}{} & \multicolumn{3}{c}{Gender} & Total \\ \cline{3-6}
 & \multicolumn{1}{c}{} & \begin{tabular}[c]{@{}c@{}}Male \\ (f)\end{tabular} & \begin{tabular}[c]{@{}c@{}}Female\\ (f)\end{tabular} &  \begin{tabular}[c]{@{}c@{}}PNTS\\ (f)\end{tabular} & \\ \cline{2-6}
\multirow{4}{*}{\begin{tabular}[c]{@{}l@{}}Post-covid\\ changes\end{tabular}} & \begin{tabular}[c]{@{}l@{}}Yes, I use a safer messaging \\ app now\end{tabular} & 20 & 20 & 0 & 40 \\
 & \begin{tabular}[c]{@{}l@{}}Yes, I didn’t use any \\ messaging app before, but \\ do use it now\end{tabular} & 20 & 23 & 1 & 44 \\
 & \begin{tabular}[c]{@{}l@{}}No, I did use it before and \\ use it now\end{tabular} & 114 & 47 & 4 & 165 \\
 & \begin{tabular}[c]{@{}l@{}}No, I didn’t use it before and \\ don’t use it now\end{tabular} & 18 & 6 & 0 & 24 \\
 & Other & 7 & 2 & 0 & 9 \\
Total &  & 179 & 98 & 5 & 282 \\ \hline
\end{tabular}
\vspace{0.8em}
\caption{Gender * post-covid changes crosstabulation}
\label{tab:gender*covid}
\end{table}
%
\begin{table}[h!tbp]
\centering
\resizebox{\textwidth}{!}{
\begin{tabular}{llcccccccc}
\hline
 & \multicolumn{1}{c}{} & \multicolumn{7}{c}{Disciplines} & Total \\ \cline{3-10}
 & \multicolumn{1}{c}{} & \begin{tabular}[c]{@{}c@{}}Engi- \\ neering(f)\end{tabular} & \begin{tabular}[c]{@{}c@{}}Social \\ Sciences (f)\end{tabular} &  \begin{tabular}[c]{@{}c@{}}Sciences\\ (f)\end{tabular} & \begin{tabular}[c]{@{}c@{}}Bio-\\ medicine (f)\end{tabular} & \begin{tabular}[c]{@{}c@{}}Human-\\ ities(f)\end{tabular} & \begin{tabular}[c]{@{}c@{}}Technol-\\ ogy(f)\end{tabular} &  \begin{tabular}[c]{@{}c@{}}Other\\ (f)\end{tabular} & \\ \cline{2-10}
\multirow{4}{*}{\begin{tabular}[c]{@{}l@{}}Post-covid\\ changes\end{tabular}} & \begin{tabular}[c]{@{}l@{}}Yes, I use a safer \\ messaging app now\end{tabular} & 7 & 11 & 5 & 5 & 1 & 10 & 1 & 40 \\
 & \begin{tabular}[c]{@{}l@{}}Yes, I didn’t use any \\ messaging app before, \\ but do use it now\end{tabular} & 13 & 12 & 5 &  4 &  2 &  7 & 1 & 44 \\
 & \begin{tabular}[c]{@{}l@{}}No, I did use it before \\ and use it now\end{tabular} &60 &  26 &  20 &  4 & 0 & 53 &  2 &  165\\
 & \begin{tabular}[c]{@{}l@{}}No, I didn’t use it \\ before and don’t \\ use it now\end{tabular} & 14 &  4 & 1 & 2 & 1 & 1 & 1 & 24 \\
 & Other & 5  & 0 & 1 & 1 & 1 & 1 & 0 & 9 \\
Total & & 99 &  53&  32&  16&  5&  72&  5&  282\\ \hline
\end{tabular}}
\vspace{0.8em}
\caption{Discipline * post-covid changes crosstabulation}
\label{tab:discipline*covid}
\end{table}
%
%\begin{figure}[t]
%\begin{center}
%    \vspace{2mm}
%    \begin{minipage}[t]{0.30\textwidth}
%	    \centering
% 	    \includegraphics[width=\textwidth]{figures/sector_whatsapp.pdf}
%    \end{minipage}
%    \begin{minipage}[t]{0.50\textwidth}
%	    \centering
% 	    \includegraphics[width=\textwidth]{figures/discipline_whatsapp.pdf}
%    \end{minipage}
%
%    \vspace{2mm}
%    \begin{minipage}[t]{0.30\textwidth}
%	    \centering
% 	    \includegraphics[width=\textwidth]{figures/sector_telegram.pdf}
%    \end{minipage}
%    \begin{minipage}[t]{0.50\textwidth}
%	    \centering
% 	    \includegraphics[width=\textwidth]{figures/discipline_telegram.pdf}
%    \end{minipage}
%\end{center}
%    \caption{Differences in the use of chatbots in class between vocational and university teaching and distributions by discipline.}
%\label{figure:chatbots_use_sector_discip}
%\end{figure}
%
%A chi-square test of independence was performed to examine the relation between instructors’ discipline, sector, and gender and any potential changes that occurred to the use of messaging apps due to the covid-19 pandemic. The relation between the latter variable and the instructors’ sector was significant, $X^{2}_{(4, N = 282)} = 9.598$, $p =0.048$. The frequencies cross tabulated are given in Table \ref{tab:sector*covid}. In effect, this finding indicates that how teachers responded to the use of messaging apps during the pandemic and in particular, whether they changed their habits in the use of apps for teaching purposes, was related to their sector (university vs vocational).
%The relationships between changes in the use of messaging apps, due to the covid-19 pandemic, and the instructors’ discipline, $X^{2}_{(24, N = 282)} = 44.856$, $p =0.006$, as well as the gender, $X^{2}_{(8, N = 282)} = 16.249$, $p= 0.039$ were also significant.

% Olia - TODO: check for repetitions in the following paragraph that you added. thank you.
Finally, we took a closer look at the post-covid changes and the variables
gender, sector, and discipline for which the chi-square test is statistically
significant as shown in Figures \ref{figure:gender*postcovid_chisq},
\ref{figure:sector*postcovid_chisq}, \ref{figure:discipline*postcovid_chisq}
respectively. Each figure shows on the left, a graph that represents the
Pearson's residuals of the chi-square test results and a table on the right part
that shows the contribution of each cell to the test. In the Pearson's residual
graphs, \textcolor{black}{the responses to the question about the use of messaging apps were shortened for the sake of clarity according to}: 1)
\textit{Did/do} stands for \textit{Yes, I did use it before and do use it now};
2) \textit{Didn't/do} for \textit{Yes, I didn’t use any messaging app before,
  but do use it now}; 3) \textit{safer} for \textit{Yes, I use a safer messaging
  app now}; 4) \textit{Didn't/Don't} for \textit{No, I didn’t use it before and
  don’t use it now}.
%
\begin{figure}[h!tbp]
\begin{center}
    \begin{minipage}[c]{0.45\textwidth}
        \centering
<<gender_postcovid_chisq, echo=FALSE, fig.width=17/2.54, fig.height=10/2.54>>=
df <-select(survey.pilot.postCOVID, PostCOVID.changes, Gender, Sector, Experience, Discipline)
df[as.character(df$Gender)!="PNTS", ] -> df1
df1[as.character(df1$Gender)!="No", ] -> df1
df1[as.character(df1$Discipline)!="No", ] -> df1
df1[as.character(df1$Experience)!="No", ] -> df1
df1[as.character(df1$Sector)!="No", ] -> df1
survey.pilot.postCovidChisq<- droplevels(df1)

covidChanges.vs.Gender<-table(survey.pilot.postCovidChisq$Gender, survey.pilot.postCovidChisq$PostCOVID.changes)
colnames(covidChanges.vs.Gender)<- c("Did/do", "Didn't/do","safer", "Didn't/don't")
covidChanges.vs.Gender<-covidChanges.vs.Gender[,c(3,2,1,4)]
chisq<-chisq.test(covidChanges.vs.Gender)
corrplot(chisq$residuals, is.corr = FALSE)
@
    \end{minipage}
    %This tables comes from (TODO: generate them automatically from the data)
    %contrib <- 100*chisq$residuals^2/chisq$statistic
    %round(contrib, 2)
    %round(chisq$stdres, 3)
    \begin{minipage}[c]{0.50\textwidth}
        \centering
    \resizebox{\textwidth}{!}{%
\begin{tabular}{llllll}
\hline
 &  & \multicolumn{4}{c}{Gender} \\ \cline{3-6}
 &  & \multicolumn{2}{c}{Female} & \multicolumn{2}{c}{Male} \\ \cline{3-6}
 &  & \multicolumn{1}{c}{\begin{tabular}[c]{@{}c@{}}Stand. \\ Residuals\end{tabular}} & \multicolumn{1}{c}{\begin{tabular}[c]{@{}c@{}}Contri-\\ bution\end{tabular}} & \multicolumn{1}{c}{\begin{tabular}[c]{@{}c@{}}Stand. \\ Residual\end{tabular}} & \multicolumn{1}{c}{\begin{tabular}[c]{@{}c@{}}Contri-\\ bution\end{tabular}} \\ \cline{2-6}
\begin{tabular}[c]{@{}l@{}}Post-\\ covid\\ changes\end{tabular} & \begin{tabular}[c]{@{}l@{}}Yes, I use a \\ safer messaging \\ app now\end{tabular} & 1.832 & 13.49\% & -1.832 & 7.50\% \\
 & \begin{tabular}[c]{@{}l@{}}Yes, I didn’t use \\ any messaging \\ app before, but\\ do use it now\end{tabular} & \textbf{2.806} & \textbf{31.21\%} & \textbf{-2.806} & \textbf{17.36\%} \\
 & \begin{tabular}[c]{@{}l@{}}No, I did use it \\ before and\\ use it now\end{tabular} & \textbf{2.850} & 15.15\% & \textbf{-2.850} & 8.43\% \\
 & \begin{tabular}[c]{@{}l@{}}No, I didn’t use \\ it before and\\ don’t use it now\end{tabular} & -1.011 & 4.40\% & 1.011 & 2.45\% \\ \hline
\end{tabular}
    %\begin{tabular}{l|r|r|r|r|}
    %\cline{2-5}
    %                             & Did/Do & Didn't/Do & Safer  & Didn' t/Don't \\ \hline
    %\multicolumn{1}{|l|}{Female} & \textbf{-2.850} & \textbf{2.806} & 1.832  & -1.011 \\
    %\multicolumn{1}{|l|}{} & 15.15\% & \textbf{31.21\%} &  13.49\% & 4.40\% \\ \hline
    %\multicolumn{1}{|l|}{Male}   & \textbf{2.850} & \textbf{-2.806} & -1.832 & 1.011 \\
    %\multicolumn{1}{|l|}{} &  8.43\% &  \textbf{17.36\%} & 7.50\% &  2.45\% \\ \hline
    %\end{tabular}%
    }
    \end{minipage}
\end{center}
    \caption{Analysis for gender and post-covid changes: left) Pearson's residuals for the Chi-square test; right) Gender*post-covid changes crosstabulation, including standardized residuals \cite{agresti2013categorical} (also adjusted standardized residuals) and contribution percentage to the total Chi-square test of each cell.}
\label{figure:gender*postcovid_chisq}
\end{figure}

As mentioned, the graphs on the left hand side represent the Pearson's
residuals, showing the difference between the observed and the expected values
for each cell. Thus, large residuals indicate that variables are not truly
independent. In our case, the blue shows positive contributions and red negative
contributions to the test. Moreover, the saturation shows how large the
contribution is in contrast to the expected value by chance. Additionally,
tables on the right side show adjusted standardized residuals that according to
\cite{agresti2013categorical}, if greater than +/-2 for cases with few cells,
indicate lack of fit of $H_0$ (in boldface in our table). Complementary, the
table also shows the percentage of contribution to the test of each cell
(highest percentage per column also in boldface).

Regarding gender in Fig. \ref{figure:gender*postcovid_chisq}, the results show a
high positive contribution from female educators that did not use a messaging
app before the COVID-19 pandemic but do use it now; approximately 31\% of the
test results are explained by this cell. In contrast, male educators that
respond the same were less than expected (negative contribution that explains
17\% of the test results). Also, it is important to highlight the female
educators that kept using these apps before and after the pandemic and
correspondingly the negative contribution of male educators to the same case
(less than expected by chance).

\begin{figure}[h!tbp]
\begin{center}
    \begin{minipage}[c]{0.45\textwidth}
        \centering
<<sector_postcovid_chisq, echo=FALSE, fig.width=17/2.54, fig.height=10/2.54, warning=FALSE>>=
covidChanges.vs.Sector<-table(survey.pilot.postCovidChisq$Sector, survey.pilot.postCovidChisq$PostCOVID.changes)
colnames(covidChanges.vs.Sector)<- c("Did/do", "Didn't/do","safer", "Didn't/don't")
covidChanges.vs.Sector<-covidChanges.vs.Sector[,c(3,2,1,4)]
chisq<-chisq.test(covidChanges.vs.Sector)
corrplot(chisq$residuals, is.corr = FALSE)
@
    \end{minipage}
    \begin{minipage}[c]{0.50\textwidth}
        \centering
    \resizebox{\textwidth}{!}{%
    %\begin{tabular}{l|r|r|r|r|}
    %\cline{2-5}
    %                             & Did/Do & Didn't/Do & Safer  & Didn' t/Don't \\ \hline
    %\multicolumn{1}{|l|}{University} & \textbf{-2.122} & 0.454 & 0.115  &  \textbf{2.939}\\
    %\multicolumn{1}{|l|}{} & 5.71\% & 0.56\% & 0.04\%  &  \textbf{25.26\%}\\ \hline
    %\multicolumn{1}{|l|}{Vocational} & \textbf{2.122} & -0.454 & -0.115 & \textbf{-2.939}\\
    %\multicolumn{1}{|l|}{} & 12.38\% & 1.21\% & 0.08\% & \textbf{54.78\%}\\ \hline
    %\end{tabular}%
    \begin{tabular}{llllll}
          \hline
           &  & \multicolumn{4}{c}{Sector} \\ \cline{3-6}
           &  & \multicolumn{2}{c}{University} & \multicolumn{2}{c}{Vocational} \\ \cline{3-6}
           &  & \multicolumn{1}{c}{\begin{tabular}[c]{@{}c@{}}Stand. \\ Residuals\end{tabular}} & \multicolumn{1}{c}{\begin{tabular}[c]{@{}c@{}}Contri-\\ bution\end{tabular}} & \multicolumn{1}{c}{\begin{tabular}[c]{@{}c@{}}Stand. \\ Residual\end{tabular}} & \multicolumn{1}{c}{\begin{tabular}[c]{@{}c@{}}Contri-\\ bution\end{tabular}} \\ \cline{2-6}
          \begin{tabular}[c]{@{}l@{}}Post-\\ covid\\ changes\end{tabular} & \begin{tabular}[c]{@{}l@{}}Yes, I use a \\ safer messaging \\ app now\end{tabular} & 0.115 & 0.04\% & -0.115 & 0.08\% \\
           & \begin{tabular}[c]{@{}l@{}}Yes, I didn’t use \\ any messaging \\ app before, but\\ do use it now\end{tabular} & 0.454 & 0.56\% & -0.454 & 1.21\% \\
           & \begin{tabular}[c]{@{}l@{}}No, I did use it \\ before and\\ use it now\end{tabular} & \textbf{-2.122} & 5.71\% & \textbf{2.122} & 12.38\% \\
           & \begin{tabular}[c]{@{}l@{}}No, I didn’t use \\ it before and\\ don’t use it now\end{tabular} & \textbf{2.939} & \textbf{25.26\%} & \textbf{-2.939} & \textbf{54.78\%} \\ \hline
          \end{tabular}%
    }
    \end{minipage}
\end{center}
    \caption{Analysis for sector and post-covid changes: left) Pearson's residuals for the Chi-square test; right) Sector*post-covid changes crosstabulation, including standardized residuals \cite{agresti2013categorical} (also adjusted standardized  residuals) and contribution percentage to the total Chi-square test of each cell.}
\label{figure:sector*postcovid_chisq}
\end{figure}
%
With respect to the sector, the results in Fig. \ref{figure:sector*postcovid_chisq} show a positive contribution from university educators being reluctant to use any messaging apps before or after the pandemic: the positive contribution of their responses explain 25\% of the test results with an adjusted residual of 2.94. On the contrary, vocational teachers were more open to it, showing a negative Pearson's residual. Also, we find larger values than expected in vocational teachers that kept using these apps after the pandemic.

\begin{figure}[h!tbp]
\begin{center}
    \begin{minipage}[c]{0.45\textwidth}
        \centering
<<discipline_postcovid_chisq, echo=FALSE, fig.width=17/2.54, fig.height=10/2.54, warning=FALSE>>=
survey.pilot.postCovidChisq %>%mutate(Discipline = ifelse(as.character(Discipline) == "No", "Hum/Social Sciences", as.character(Discipline))) %>%
  mutate(Discipline = ifelse(as.character(Discipline) == "Other", "Hum/Social Sciences", as.character(Discipline))) %>%
  mutate(Discipline = ifelse(as.character(Discipline) == "Engineering", "Tech/Engineering", as.character(Discipline))) %>%
  mutate(Discipline = ifelse(as.character(Discipline) == "Technology", "Tech/Engineering", as.character(Discipline))) %>%
  mutate(Discipline = ifelse(as.character(Discipline) == "Humanities", "Hum/Social Sciences", as.character(Discipline))) %>%
  mutate(Discipline = ifelse(as.character(Discipline) == "Social Sciences", "Hum/Social Sciences", as.character(Discipline))) %>%
  mutate(Discipline = ifelse(as.character(Discipline) == "Sciences", "Biomed/Sciences", as.character(Discipline))) %>%
  mutate(Discipline = ifelse(as.character(Discipline) == "Biomedicine", "Biomed/Sciences", as.character(Discipline)))->survey.pilot.postCovidChisq
covidChanges.vs.Discipline<-table(survey.pilot.postCovidChisq$Discipline, survey.pilot.postCovidChisq$PostCOVID.changes)
colnames(covidChanges.vs.Discipline)<- c("Did/do", "Didn't/do","safer", "Didn't/don't")
covidChanges.vs.Discipline<-covidChanges.vs.Discipline[,c(3,2,1,4)]
chisq<-chisq.test(covidChanges.vs.Discipline)
corrplot(chisq$residuals, is.corr = FALSE)
@
    \end{minipage}
    \begin{minipage}[c]{0.50\textwidth}
        \centering
    \resizebox{\textwidth}{!}{%
\begin{tabular}{llllllll}
\hline
 &  & \multicolumn{6}{c}{Discipline} \\ \cline{3-8}
 &  & \multicolumn{2}{c}{Biomed/Sciences} & \multicolumn{2}{c}{Hum/Social} & \multicolumn{2}{l}{Tech/Engineering} \\ \cline{3-8}
 &  & \multicolumn{1}{c}{\begin{tabular}[c]{@{}c@{}}Stand. \\ Residuals\end{tabular}} & \multicolumn{1}{c}{\begin{tabular}[c]{@{}c@{}}Contri-\\ bution\end{tabular}} & \multicolumn{1}{c}{\begin{tabular}[c]{@{}c@{}}Stand. \\ Residual\end{tabular}} & \multicolumn{1}{c}{\begin{tabular}[c]{@{}c@{}}Contri-\\ bution\end{tabular}} & \begin{tabular}[c]{@{}l@{}}Stand. \\ Residuals\end{tabular} & \begin{tabular}[c]{@{}l@{}}Contri-\\ bution\end{tabular} \\ \cline{2-8}
\begin{tabular}[c]{@{}l@{}}Post-\\ covid\\ changes\end{tabular} & \begin{tabular}[c]{@{}l@{}}Yes, I use a \\ safer mes-\\ saging app \\ now\end{tabular} & 1.452 & \textbf{9.94\%} & 1.566 & 11.04\% & \textbf{-2.448} & \textbf{13.29\%} \\
 & \begin{tabular}[c]{@{}l@{}}Yes, I didn’t \\ use any mes-\\ saging app \\ before, but\\ do use it now\end{tabular} & 0.733 & 2.50\% & \textbf{2.079} & \textbf{19.19\%} & -2.318 & 11.76\% \\
 & \begin{tabular}[c]{@{}l@{}}No, I did use \\ it before and\\ use it now\end{tabular} & -1.265 & 3.50\% & \textbf{-2.729} & 15.55\% & \textbf{3.278} & 11.06\% \\
 & \begin{tabular}[c]{@{}l@{}}No, I didn’t \\ use it before \\ and don’t use \\ it now\end{tabular} & -0.588 & 1.75\% & 0.055 & 0.01\% & 0.412 & 0.40\% \\ \hline
\end{tabular}
    %\begin{tabular}{l|r|r|r|r|}
    %\cline{2-5}
    %                             & Did/Do & Didn't/Do & Safer & Didn' t/Don't \\ \hline
    %\multicolumn{1}{|l|}{Biomed/Sciences} & -1.265 &  0.733 & 1.452 & -0.588 \\
    %\multicolumn{1}{|l|}{} & 3.50\% &  2.50\% & 9.94\% & 1.75\% \\ \hline
    %\multicolumn{1}{|l|}{Hum/Social} &  \textbf{-2.729} & \textbf{2.079} &  1.566 & 0.055\\
    %\multicolumn{1}{|l|}{} &  15.55\% & \textbf{19.19\%} &  11.04\% & 0.01\%\\ \hline
    %\multicolumn{1}{|l|}{Tech/Engineering} &  \textbf{3.278} &  \textbf{-2.318} &  \textbf{-2.448} & 0.412\\
    %\multicolumn{1}{|l|}{} &  11.06\% &  11.76\% &  \textbf{13.29\%} & 0.40\%\\ \hline
    %\end{tabular}%
    }
    \end{minipage}
\end{center}
    \caption{Analysis for discipline and post-covid changes: left) Pearson's residuals for the Chi-square test; right) Discipline*post-covid changes crosstabulation, including standardized residuals \cite{agresti2013categorical} (also adjusted standardized residuals) and contribution percentage to the total Chi-square test of each cell.}
\label{figure:discipline*postcovid_chisq}
\end{figure}
%
Finally, statistical significance was found for the discipline and post-covid
changes. Regarding disciplines (see
Fig. \ref{figure:discipline*postcovid_chisq}), the number of educators from
Humanities and Social Sciences that started using messaging apps after the
pandemic is larger than expected and these number of responses explains about
19\% of the chi-square test results. Also, the number of educators from
Technology or Engineering that started using safer alternatives \footnote{This addresses the issue of insecure/unsafe applications which do not encrypt messages or can be easily compromised exposing information} are less than
expected (compared to the educators from other disciplines). Bear in mind that
some disciplines were merged to avoid very low number of responses for some of
the cells.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% RQ2 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{RQ2 - Which kind of chatbots would teachers find useful in their classes?}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

For answering RQ2, teachers were provided with a list of different potential chatbot functionalities (use cases) and were requested to respond on whether each given use case would be useful in their classes. The findings are summarized in Table \ref{tab:chatbottypes}.

\begin{table}[h!tbp]
\begin{tabular}{lllll}
\hline
\multicolumn{1}{c}{Chatbot use cases} & \multicolumn{2}{c}{Yes} & \multicolumn{2}{l}{No} \\ \cline{2-5}
\multicolumn{1}{c}{} & \multicolumn{1}{c}{Frequency} & \multicolumn{1}{c}{\%} & Frequency & \multicolumn{1}{c}{\%} \\ \hline
Answering to students' FAQs & 148 & 52.5 & 134 & 47.5 \\
Assigning student grades & 113 & 40.1 & 169 & 59.9 \\
Facilitating agenda information & 171 & 60.6 & 111 & 39.4 \\
Sharing class materials & 136 & 48.2 & 146 & 51.8 \\
Others & 25 & 8.9 & 257 & 91.1 \\ \hline
\end{tabular}
\caption{Perceived useful chatbot use cases}
\label{tab:chatbottypes}
\end{table}

The most favorable use case for chatbots in
Higher Education and vocational training is their use for the facilitation of an
agenda formation (171 positive response, 60.6\%), followed by the FAQs use case
(148 positive responses, 52.5\%) and the sharing class material use case (136;
48.2\%). A chi-square test of independence was performed to examine the relation
among participants’ preferences for particular chatbot use cases. Out of the 171 teachers who consider useful the use of
chatbots for agenda preparation in the class, 103 also consider useful chatbots’
use for FAQs. The relation between agenda and the FAQs use case was significant, $X^{2}_{(1, N = 282)} =
10.467$, $p =0.001$. The frequencies cross tabulated are given in Table
\ref{tab:agenda*faq}.

\begin{table}[h!tbp]
\begin{tabular}{llccc}
\hline
 & \multicolumn{1}{c}{} & \multicolumn{2}{c}{FAQs} & Total \\ \cline{3-5}
 & \multicolumn{1}{c}{} & Yes & No &  \\ \cline{2-5}
\multirow{2}{*}{Agenda} & Yes & 103 & 68 & 171 \\
 & No & 45 & 66 & 111 \\
Total &  & 148 & 134 & 282 \\ \hline
\end{tabular}
\vspace{0.8em}
\caption{Agenda use case * FAQs use case crosstabulation }
\label{tab:agenda*faq}
\end{table}

Answers to these questions are plotted in
Figs. \ref{figure:gender*chatbots_types}, \ref{figure:sector*chatbots_types},
\ref{figure:experience*chatbots_types}, \ref{figure:disc*chatbots_types} grouped
by gender, sector, years of experience in education, and discipline
respectively. As it was a multiple choice question, the counts are over the
total number of people that answered.

\begin{figure}[h!tbp]
\begin{center}
    \begin{minipage}[t]{0.85\textwidth}
        \centering
<<gender_chatbottype, echo=FALSE, fig.width=19/2.54, fig.height=5.5/2.54>>=
survey.pilot.ChatbotType <- within(survey.pilot.ChatbotType, ChatbotType <- factor(ChatbotType, levels=names(sort(table(ChatbotType), decreasing=TRUE))))
survey.pilot.ChatbotType[as.character(survey.pilot.ChatbotType$ChatbotType)!="OtherBot", ] -> survey.pilot.ChatbotType
#survey.pilot.ChatbotType[as.character(survey.pilot.ChatbotType$Gender)!="PNTS", ] -> survey.pilot.ChatbotType
ggplot(survey.pilot.ChatbotType, aes( x = ChatbotType, y = ..count.., group=Gender, fill = Gender ) ) + geom_bar()+ theme(axis.text.x = element_text(angle = 90)) +coord_flip()+ theme_tufte()+ theme(axis.title.y=element_blank())
@
%       \includegraphics[width=\textwidth]{figures/gender_chatbottype.pdf}
        \vspace{-0.5cm}
        \caption{Count of types of chatbots for class perceived as the most useful for teachers grouped by gender (PNTS stands for Prefer Not To Say).}
        \label{figure:gender*chatbots_types}
    \end{minipage}
    \begin{minipage}[t]{0.85\textwidth}
        \centering
<<sector_chatbottype, echo=FALSE, fig.width=19/2.54, fig.height=5.5/2.54>>=
ggplot(survey.pilot.ChatbotType, aes( x = ChatbotType, y = ..count.., group=Sector, fill = Sector ) ) + geom_bar()+ theme(axis.text.x = element_text(angle = 90)) +coord_flip()+ theme_tufte()+ theme(axis.title.y=element_blank())
@
%       \includegraphics[width=\textwidth]{figures/sector_chatbottype.pdf}
        \vspace{-0.5cm}
        \caption{Count of types of chatbots for class perceived as the most useful for teachers grouped by sector.}
        \label{figure:sector*chatbots_types}
    \end{minipage}
    \begin{minipage}[t]{0.85\textwidth}
        \centering
<<experience_chatbottype, echo=FALSE, fig.width=19/2.54, fig.height=5.5/2.54>>=
ggplot(survey.pilot.ChatbotType, aes( x = ChatbotType, y = ..count.., group=Experience, fill = Experience ) ) + geom_bar()+ theme(axis.text.x = element_text(angle = 90)) +coord_flip()+ theme_tufte()+ theme(axis.title.y=element_blank())
@
%       \includegraphics[width=\textwidth]{figures/experience_chatbottype.pdf}
        \vspace{-0.5cm}
        \caption{Count of types of chatbots for class perceived as the most useful for teachers grouped by experience.}
        \label{figure:experience*chatbots_types}
    \end{minipage}
    \begin{minipage}[t]{0.85\textwidth}
        \centering
<<discipline_chatbottype, echo=FALSE, fig.width=19/2.54, fig.height=5.5/2.54>>=
survey.pilot.ChatbotType %>%
  mutate(Discipline = ifelse(as.character(Discipline) == "Engineering", "Tech/Engineering", as.character(Discipline))) %>%
  mutate(Discipline = ifelse(as.character(Discipline) == "Technology", "Tech/Engineering", as.character(Discipline)))->survey.pilot.ChatbotType
survey.pilot.ChatbotType[as.character(survey.pilot.ChatbotType$Discipline)!="No", ] -> survey.pilot.ChatbotType
ggplot(survey.pilot.ChatbotType, aes( x = ChatbotType, y = ..count.., group=Discipline, fill = Discipline ) ) + geom_bar()+ theme(axis.text.x = element_text(angle = 90)) +coord_flip()+ theme_tufte()+ theme(axis.title.y=element_blank())
@
%       \includegraphics[width=\textwidth]{figures/experience_chatbottype.pdf}
        \vspace{-0.5cm}
        \caption{Count of types of chatbots for class perceived as the most useful for teachers grouped by discipline.}
        \label{figure:disc*chatbots_types}
    \end{minipage}
\end{center}
%    \caption{Counts of chatbots to be used in class perceived as the most useful for teachers.}
%\label{figure:chatbots_types}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% RQ3 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{RQ3 - Which kind of interaction do teachers prefer with their students?}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{figure}[h!tbp]
\centering
<<messaging_organization, echo=FALSE, fig.width=19/2.54, fig.height=5/2.54>>=
survey.pilot.MessagingOrganization %>%mutate(Messaging.Organization = ifelse(as.character(Messaging.Organization) == "A chat with all the students and teachers in my Faculty/School", "A chat with all the students and teachers in my Faculty/School/all years", as.character(Messaging.Organization))) %>%
  mutate(Messaging.Organization = ifelse(as.character(Messaging.Organization) == "A chat with all the students and teachers in the same studies (all years)", "A chat with all the students and teachers in my Faculty/School/all years", as.character(Messaging.Organization)))->df2

df2 %>%mutate(Messaging.Organization = ifelse(as.character(Messaging.Organization) == "A chat with all the students and teachers in my Faculty/School/all years", "Chat interactions among all students, teachers of the School/Faculty", as.character(Messaging.Organization))) %>%
  mutate(Messaging.Organization = ifelse(as.character(Messaging.Organization) == "I'd rather leave students self-organize and not be part of the chat", "Teacher not being part of the interaction", as.character(Messaging.Organization)))%>%
  mutate(Messaging.Organization = ifelse(as.character(Messaging.Organization) == "A chat with my students in the same course", "Chat interactions among students in the same course", as.character(Messaging.Organization)))%>%
  mutate(Messaging.Organization = ifelse(as.character(Messaging.Organization) == "A chat with all students and teachers in the same semester/year", "Chat interactions among students of the same study year and teachers", as.character(Messaging.Organization)))->df2

df2 <- within(df2, Messaging.Organization <- factor(Messaging.Organization, levels=names(sort(table(Messaging.Organization), decreasing=TRUE))))
ggplot(df2, aes( x = Messaging.Organization, y = ..count.., group=Sector, fill = Sector ) ) + geom_bar()+ theme(axis.text.x = element_text(angle = 90))+coord_flip()+ theme(axis.title.y=element_blank())
@
%\includegraphics[width=\textwidth]{figures/messaging_organization.pdf}
\caption{Distribution of teachers' preferences for the chat groups with their students: from groups only with their students from a specific course to groups with greater social interaction with all students in their School or Faculty.}
\label{figure:chatbots_messagingorganization}
\end{figure}

Since chatbots are intended to mediate or help in this interaction, it is
essential to understand the kind of interaction teachers prefer. Chatbots should
address those modes, and not others. In order to find out these modes, and thus
answering RQ3, teachers were provided with a list of different kind of
interactions that may take place among students and between students and the
teacher, with the use of messaging apps. The findings are summarized in Table
\ref{tab:kindinteractions}.

\begin{table}[h!tbp]
\begin{tabular}{lllll}
\hline
\multicolumn{1}{c}{Kind of interactions} & \multicolumn{2}{c}{Yes} & \multicolumn{2}{l}{No} \\ \cline{2-5}
\multicolumn{1}{c}{} & \multicolumn{1}{c}{Frequency} & \multicolumn{1}{c}{\%} & Frequency & \multicolumn{1}{c}{\%} \\ \hline
%\begin{tabular}[c]{@{}l@{}}Chat interactions among \\ students in the same course\end{tabular} & 171 & 60.6 & 110 & 39.0 \\
%\begin{tabular}[c]{@{}l@{}}Chat interactions among students, \\ teachers of the School/Faculty\end{tabular} & 13 & 4.6 & 269 & 95.4 \\
%\begin{tabular}[c]{@{}l@{}}Chat interactions among students \\ of the same study year and teachers\end{tabular} & 51 & 18.1 & 231 & 81.9 \\
%\begin{tabular}[c]{@{}l@{}}Teacher not being part of the \\ interaction\end{tabular} & 113 & 40.1 & 169 & 59.9 \\ \hline
\begin{tabular}[c]{@{}l@{}}Chat interactions among \\ students in the same course\end{tabular} & 166 & 59.1 & 115& 40.9 \\
\begin{tabular}[c]{@{}l@{}}Chat interactions among students, \\ teachers of the School/Faculty\end{tabular} & 19 & 6.7 & 266 & 94.3 \\
\begin{tabular}[c]{@{}l@{}}Chat interactions among students \\ of the same study year and teachers\end{tabular} & 58 & 20.5 & 224 & 79.4 \\
\begin{tabular}[c]{@{}l@{}}Teacher not being part of the \\ interaction\end{tabular} & 81 & 28.7 & 201 & 71.3 \\ \hline
\end{tabular}
\vspace{0.8em}
\caption{Kind of interactions preferred}
\label{tab:kindinteractions}
\end{table}

As it can be seen by the answer to the first and last question, in general
teachers do not want to participate in a chat group with students; either they
want to simply leave the students alone in their own chat group, or otherwise
they prefer not to be part of that interaction. In general, that is going to be
the case no matter what; it is well known that students organize their own chat
groups with many (and not always conveyable) intentions, so teachers do not want
to take any part in these informal or non institutionally-supported chat
groups. Overwhelmingly, they do not want to participate in these kind of chat
groups with students, but even less so if it includes the rest of the faculty.

The answer, then, to this research question, is that, in general, teachers do
{\em not} want to have interaction with students in a chat group. To a certain
point, this would seem to contradict results \cite{gachago2015crossing}, although this might be due to cultural
attitudes, or other factors such as the average size of classes. It does
confirm, however, that the challenges cited in that study, and possibly others,
are an obstacle to the adoption of mobile (and other) instant messaging among
the community of surveyed teachers.

Regarding the social factor of chat groups that teachers use in class, according
to Fig. \ref{figure:chatbots_messagingorganization} the vast majority prefer
small groups only with the students from the same course. These are more focused
groups with specific goals and dedicated to the organization of the course and
its tasks, and from the pedagogic point of view \textcolor{black}{it also seems more appropriate to improve
the learning process}. Interestingly, and as a cross-check of the answers above,
about 30\% of teachers consider they should not be part of the chat group. This
might seem to contradict the results of the other survey, but in fact, being a
result of different surveys, to a certain point affirms the same thing: there is
a great amount of teachers that would be against being in a chat group with
students. \textcolor{black}{However, students will still need to get the services
that the university, through chatbots, provide, thus opening the door for deployment of chatbots
without the intervention of the professor, simply tapping university provided
services \cite{bernier2002services}.}
Whether that's a slim majority or not, that's open to debate (and
would probably need a more pointed survey focused on this), but what is clear is
that forcing teachers to create chat groups with students and participate in
them would create a certain amount of resistance. Also, only university teachers
find interesting a group with all the students and teachers in their own Faculty
or School. The lack of teachers from vocational education here may be the
consequence of using specific language such as ``Faculty''. Moreover, the fact
that many universities and schools already use these groups for administrative
and social interaction (e.g. \cite{dibitonto2018chatbot}) might be the reason
for the low percentage of teachers that selected this response.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% RQ4 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\subsection{RQ4 - What kind of interaction media features do teachers value the most?}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%A total of 205 teachers responded to the second survey, 124 males (60.5\%), 65 females (31.7\%) and 16 individuals who chose not to identify their gender (7.8\%). The majority of the teachers were aged 45-55 years old (70; 34.1\%), followed by individuals aged 35-45 (67, 32.7\%), and 25-35 (42; 20.5\%), while the minority was aged >55 years old (26; 12.7\%). In terms of teaching experience, 59 teachers (28.8\%) had an experience of 16-25 years, followed by teachers who had a teaching experience of 06-15 years (51; 24.9\%), 0-5 years (51; 24.9\%) and more than 25 years of experience (44; 21.5\%). The teachers were asked about what kind of interaction media features they value the most; the frequencies and percentages of their responses are given in the table that follows.

\begin{figure}[h!tbp]
\begin{center}
    \begin{minipage}[t]{0.3\textwidth}
        \centering
<<gender_synch, echo=FALSE, fig.width=10/2.54, fig.height=7/2.54>>=
survey.interaction[as.character(survey.interaction$Gender)!="PNTS", ] -> survey.interaction
ggplot(survey.interaction, aes(x = Interaction.Synchrony..or.not., group = Gender)) + geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count", show.legend = FALSE) + scale_y_continuous(labels=scales::percent) + ylab("relative frequencies") + facet_grid(~factor(Gender, levels=c('Female','Male','PNTS'))) + theme(axis.title.y=element_blank(), axis.title.x=element_blank())
@
%       \includegraphics[width=\textwidth]{figures/gender_synch.pdf}
    \end{minipage}
    \begin{minipage}[t]{0.3\textwidth}
        \centering
<<experience_sync, echo=FALSE, fig.width=10/2.54, fig.height=7/2.54>>=
ggplot(survey.interaction, aes(x = Interaction.Synchrony..or.not., group = Experience)) + geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count", show.legend = FALSE) + scale_y_continuous(labels=scales::percent) + ylab("relative frequencies") + facet_grid(~factor(Experience, levels=c("0-05 years","06-15 years", "16-25 years", "> 25 years"))) + theme(axis.title.y=element_blank(), axis.title.x=element_blank())
@
%       \includegraphics[width=\textwidth]{figures/experience_synch.pdf}
    \end{minipage}
%    \begin{minipage}[t]{0.3\textwidth}
%        \centering
%<<sector_synch, echo=FALSE, fig.width=10/2.54, fig.height=8.25/2.54>>=
%ggplot(survey.interaction, aes(x = Interaction.Synchrony..or.not., group = Sector)) + geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count", show.legend = FALSE) + scale_y_continuous(labels=scales::percent) + ylab("relative frequencies") + theme(axis.title.y=element_blank(), axis.title.x=element_blank())
%@
%       \includegraphics[width=\textwidth]{figures/sector_synch.pdf}
%    \end{minipage}
    \begin{minipage}[t]{0.6\textwidth}
        \centering
<<discipline_synch, echo=FALSE, fig.width=18.6/2.54, fig.height=7.7/2.54>>=
ggplot(survey.interaction, aes(x = Interaction.Synchrony..or.not., group = Discipline)) + geom_bar(aes(y = ..prop.., fill = factor(..x..)), stat="count", show.legend = FALSE) + scale_y_continuous(labels=scales::percent) + ylab("relative frequencies") + facet_grid(~factor(Discipline, levels=c('Biomedicine','Humanities','Social Sciences', "Sciences", "Tech/Engineering"))) + theme(axis.title.y=element_blank(), axis.title.x=element_blank())
@
%       \includegraphics[width=\textwidth]{figures/discipline_synch.pdf}
    \end{minipage}
\end{center}
    \caption{Results of teachers' communication preferences: synchronous or asynchronous interaction with students (grouped by gender, years of experience, and discipline).}
\label{figure:chatbots_synch}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% RQ4 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{RQ4 - What kind of interaction media features do teachers value the most? }

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%A total of 205 teachers responded to the second survey, 124 males (60.5\%), 65 females (31.7\%) and 16 individuals who chose not to identify their gender (7.8\%). The majority of the teachers were aged 45-55 years old (70; 34.1\%), followed by individuals aged 35-45 (67, 32.7\%), and 25-35 (42; 20.5\%), while the minority was aged $>$~55 years old (26; 12.7\%). In terms of teaching experience, 59 teachers (28.8\%) had an experience of 16-25 years, followed by teachers who had a teaching experience of 06-15 years (51; 24.9\%), 0-5 years (51; 24.9\%) and more than 25 years of experience (44; 21.5\%). The teachers were asked about what kind of interaction media features they value the most \textcolor{black}{; i.e. what their preferred features are in the applications/tools that they use to interact with their students.} The frequencies and percentages of their responses are given in Table \ref{tab:interaction}.

\begin{table}[h!]
\begin{tabular}{lcccc}
\hline
Interaction features & \multicolumn{2}{c}{Yes} & \multicolumn{2}{c}{No} \\
\hline
                    &  Freq. & \% &  Freq. & \% \\
                    \cline{2-5}

Analytics & 108 & 52.7 & 97 & 47.3 \\
Connectivity & 119 & 58.0 & 86 & 42.0 \\
Familiarity & 121 & 59.0 & 84 & 41.0 \\
Hidden Phone Number & 113 & 55.1 & 92 & 44.9 \\
Horizontally & 134 & 65.4 & 71 & 34.6 \\
Official formation & 42 & 20.5 & 163 & 79.5 \\
Pluggability & 65 & 31.7 & 140 & 68.3 \\
Sustainability & 157 & 76.6 & 48 & 23.4\\
Unidirectionality & 27 & 13.2 & 178 & 86.8 \\
Officiality & 150 & 73.2 & 55 & 26.8 \\
Synchrony & 45 & 22.0 & 160 & 78.0\\
Other & 5 & 2.4 & 200 & 97.6 \\
\hline
\end{tabular}
\vspace{0.8em}
\caption{Interaction media features valued by teachers. Details about each of these features can be read in Appendix I (Section \ref{app:questions}, second survey).}
\label{tab:interaction}
\end{table}

\textcolor{black}{As it can be seen in the table, the bulk of responders express their wish to use a sustainable and official application, i.e. both being approved or provided by the University (or the educational center), and also maintained by the technical staff of the center instead of giving this responsibility/task to the teacher.
In addition, it is very important for the teachers that the used tools that all the members of the class communicate, including themselves; as well as being already known or used in everyday tasks, e.g. Telegram or WhatsApp.}

A chi-square test of independence was performed to examine the potential relationship among the various features. The statistically significant correlations are given in Table \ref{tab:interactionfeatures}.

\begin{table}[h!tbp]
\centering
\resizebox{\textwidth}{!}{\begin{tabular}{lllllllll}
\hline
 & Connectivity & Familiarity  & Horizontality & \begin{tabular}[c]{@{}l@{}}Official\\  formation\end{tabular} & Pluggability & Sustainability & Unidirectionality & Officiality \\ \hline
Analytics &  $12.172^{***}$ & $4.855^{*}$ & ns & $4.144^{*}$ &  $12.490^{***}$ & $4.314^{*}$ & ns & ns \\
Connectivity & - & ns & $5.971^{***}$ & ns & $16.286^{***}$ & ns & ns & ns \\
Hidden phone &  &  & ns & $5.678^{*}$ & $4.683^{*}$ & ns & $4.515^{*}$ & ns \\
Horizontality &  &  - & ns & ns & ns & ns & ns & $5.453^{*}$ \\
\begin{tabular}[c]{@{}l@{}}Official\\  formation\end{tabular} &  & &  & - & $8.163^{**}$ & ns & ns & ns  \\
\end{tabular}}
\vspace{0.8em}
\caption{Interaction media features analysis ($^{***}p<0.001$, $^{**}p<0.01$, $^{*}p<0.05$). Columns and rows with no significant interaction have been suppressed for clarity. Every cell value represents $X^{2}_{(1, N = 205)}$. Details about each feature can be read in Appendix I.}
\label{tab:interactionfeatures}
\end{table}

Finally, a two-step cluster analysis was conducted, with a log-likelihood
distance measure adopted, to explore how the use cases could be grouped based on
their preferences of specific interaction media features, and which features
have larger predictor importance for the clustering (see
Fig. \ref{figure:predictor}). Two clusters occurred from the analysis: cluster 1
(n=78, 60.9\%) is formed with instructors who did not value interaction media
pluggability, connectivity, official formation, but valued interaction media
analytics, familiarity, support by the institution; cluster 2 (n=50,
39.1\%) groups instructors that did not value media
analytics, official formation, familiarity, but valued pluggability,
connectivity, and support by the institution. Interaction media pluggability
appears to have the most important predictor importance in the clustering of
cases (predictor importance = 1.0), whilst media unidirectionality was the least
important factor (predictor importance = 0.04). The cluster quality is fair, but
not a good one (silhouette measure of cohesion and separation).

\begin{figure}[h!tbp]
\begin{center}
\includegraphics[width=\textwidth]{figures/2clusteranalysis.pdf}
\end{center}
    \caption{Two-step cluster analysis for interaction media features: predictor importance}
\label{figure:predictor}
\end{figure}

% ********************** Discussion section *******************
\section{Discussion and implications}

Our initial intention in the design of these surveys was to probe the opinions
of tertiary education teachers in the introduction of chatbot technologies in
class. We wanted to find the answer to four research questions, of which the
first was to check whether teachers already used some kind of messaging
application, and the one they preferred. Depending on the answer, different
technologies could be (or not) available for the introduction of chatbots. In
this case, the design of the survey introduced some ambiguity on what we really
consider ``messaging technology'', that was eventually resolved by the answers.
Apparently, anything that sends messages (even if they are not instant, even if
they do not have a dedicated application) was the concept in the minds of the
surveyed. In this case, we would like
to add a clarification. While we had in mind, as indicated by the possible answers,
instant messaging applications when we elaborated this question,
we included these institution-provided applications mainly for completion. But
it should be noted that, in most cases, they are not {\em instant} messaging
apps in the way Telegram or Facebook Messenger are; in most cases, they are
messaging facilities provided by learning management systems such as (usually)
Moodle. So the answer to this question must be understood in two different ways:
their preference for the {\em provider} of the technology (the university itself, any
company) as well as the {\em type} of technology (instant, app-based messaging
systems vs. web-based messaging systems integrated in another learning management system).

This inclusion of university-provided messaging applications in
the survey might explain why there are so few teachers, just 6.7\%, that answered they did not
use any messaging application. In general, universities provide official channels
to communicate with students official academic information such as the exam schedule
or the grades. The statistical mode is to use that kind of messaging. Some other
teachers might not use it, but use some extraofficial or opt-in channels such as the others included in the survey; in the end, between these official and extra-official channels, most people cover their needs to communicate with students. This matches
our subjective experience, in general.

Additionally, the question on COVID reflected the big challenge that represents
changing habits, even in the case of a major crisis. Very few responders actually
changed their habits, although comparing women vs. men as well as those teaching social
sciences show some significant differences. We do not have any kind of explanation for
this behavior, other than technologies are perceived, and adopted, in a
different way depending on background and, apparently, gender.

One of the most interesting questions arising from the analysis of the
results, and that should require further study, is: Are teachers really aware
of the possibilities of chatbots in the classroom? The survey
only listed a few functionalities for the chatbots, and just a few subjects selected
the option "Others". This may suggest that most teachers did not really understand
at this point the actual potential of chatbots, since the responses listed were by no means exhaustive. If the responders were more acquainted with the possibilities of this technology, this response would possibly have shown up many more times. Another way to look at it is that, unless you provide the specific instructions on how to use a
tool, the teaching staff is already too busy or burnt out to think about innovation.

We can speculate that teachers' opinions of chatbots are influenced by negative
experiences with customer support chatbots, which is what most people are likely to have
experienced. Further study may be needed to confirm this, but it is important to 
keep in mind when introducing chatbot technologies in the classroom.
To potentially change this existing perception, it will be necessary to make the chatbot as simple and satisfying to use as possible.

Regarding gender grouping, both genders agree in their uses, so it seems not
relevant at all. However, when we look at the plot containing the sector differentiation,
it is interesting to note that university teachers are more likely to use chatbots 
for grades and FAQs compared to other tertiary education teachers.
This seems reasonable as the frequency of seeing the students
is lower at the University so, an automatic tool for grade notification for
continuous evaluation would be more helpful.
Furthermore, the high student-to-teacher
ratios that university teachers often face could be another reason why
a higher number of them are willing to use chatbots for FAQs, as this could
reduce misunderstandings and email overload from students.

The results, when grouping by experience, confirm that
experienced teachers are not afraid of changes and are willing to accept new
technological challenges; at least, in the same way that younger teachers
are. From this concrete plot it is interesting as well to observe the fact that very few
young/inexperience teachers are willing to use chatbots for other uses. This
reinforces the hypothesis thrown at the beginning of this section that younger
teachers have less experience to consider new ways of using chatbots in their daily work.

As it appears in the responses presented above, teachers’ responses are equally
distributed between positive and negative in relation to the media features:
analytics, connectivity, and hidden phone number; that is, none of them is a
prevailing factor on choosing one technology over another. However, for the features: familiarity,
horizontally, sustainability and officiality the majority provided a positive
response, designating their preference to these features. On the other hand, the
majority of teachers seems to consider not valuable the presence of official
formation, pluggability, unidirectionality and synchrony in the interaction
media features of messaging apps.

From a qualitative point of view, these results show consensus in a few key aspects:
\begin{itemize}
 \item Teachers want to have support from their corresponding institution. This
   is probably due to the strict European data regulations \textcolor{black}{on data protection},
   but also to avoid the overhead required to \textcolor{black}{sign up new students every year.
   In general support from specific formation and access to IT help desk for specific features
   enhances the possibility of adoption by the teachers or the possibility of a successful
   adoption that improves the learning outcomes. This is supported by the result of the two
   surveys: they prefer whatever "messaging" application is provided by the university, and
   also they value sustainability as well as "officiality" of the platform, as shown in
   Table \ref{tab:interaction}}.
 \item Teachers need sustainability. The main reason is that implementing
   changes is expensive in terms of effort and prone to errors in the first
   years of implementing new procedures, therefore, it is logical that if
   chatbots or messaging platforms are to be introduced, they want this change
   to be as permanent as possible.
 \item Keep it Simple. Media interaction requirements are simple and only a 2.4\% is requesting more features. This could be a sign of technological burn out produced by the previous and current courses where, due to the pandemic, the use of computer has increased significantly.
\end{itemize}

An analysis of survey questions related to RQ4 seems to indicate that teachers
prefer to share an interaction space with their student, as
shown in Tables \ref{tab:kindinteractions} and \ref{tab:interaction}, where a majority of responders
indicate they {\em don't} want unidirectional communication; a preference of
{\em horizontal} interaction, but also the answer to the kind of
chatbots preferred that features FAQ bots prominently, which might indicate a
need to delegate or offload part
of the burden of answering, and attending, every single question posed by
students. In an horizontal setting, other students can answer if the teacher
does not do it immediately. This might also be the reason why the possibility of
using tools from which analytics can be extracted is also valued by teachers:
that way, activity in the chat tool, answering questions, regularity, can all be
observed and valued as part of the process of assessing achievement of learning
objectives by students.

% Only some COVID-response relationship.
In our survey, we could not find an answer about the right moment for introducing chatbot technologies. Response to questions related to change of
behavior during and after the first stay at home, stage of the pandemic do not suggest that a major (or minor)
crisis could be an opportunity to introduce new technologies, since it does not
bring major changes in attitude (as a minor "crisis" we could mention any institutional event that brings about methodological changes, such as new higher education laws\footnote{We seem to have one of these, at least in
  Spain, every 10 years or so.}), or simply introduction of new degrees. Although
changes driven by external forces do offer the chance of piloting new technologies, they do not
seem to bring internal changes in attitudes in the teaching staff (which, after all, is
bound to be the same). In absence of a clear answer in this direction, the right
moment to introduce new technologies is always when the IT and managing staff is
ready to support it (since, as we have seen before, "official" support is one of
the factors that is most valued by teachers).

% ********************** CONCLUSIONS ****************************

\section{Conclusions}
\label{sec:concl}

The key findings of this study shed light on educators’ preferences to use
messaging applications that their institutions support. Technology adoption
literature often \textcolor{black}{focuses} on users’ perceptions of the technology’s usefulness and
ease of use as important prerequisites for successful adoption and
utilization. Nevertheless, in higher education, institutions’ role in
integrating these tools to their educational systems can improve the uptake of
these applications and shape the social and educational experiences of their
students. \textcolor{black}{To accomplish this fruitful integration}, institutions have to ensure that
these messaging applications not only are GDPR (General Data Protection Regulation) compliant to keep students’ data
secure, but also should provide IT support to all stakeholders who use the
applications.

Comparing these results with those obtained in student surveys
in \cite{MoraChatbots2021:anon}, we have understood better the differences between the teachers’
points of view and/or intentions and that of their students when using the
messaging applications in higher education.  Teachers tend to adopt technologies
supported by their institutions.  The fact they do so could be caused by a desire to ensure that their
universities can oversee their efforts in supporting their students during the
learning process.  Another reason could be their familiarity with the technology
provided by their institutions. On the other hand, students use the
non-institution messaging applications to form informal discussion groups with
their peers.  It is worth noting that peer support and collaboration is
inseparable from learning \cite{timmis2012constant} and correlates positively
with higher retention in higher education \cite{o2014front}.  Therefore, both
perspectives are complementary and play different roles in promoting the
learning process.

As in many other methodologies that try to assess technology acceptance, answers to the survey suggest
that the introduction of simple, and institution-supported, instant messaging
and chatbot technologies would increase the perceived usefulness (which is one
of the key metrics in technology acceptance models). Since most institutions
already have some {\em virtual campus} or learning management system, adding
some easy automation, or connection to personal instant messaging tools, could
really help onboard the learning community on these new technologies.

What we are going to propose next is a possible process of
technology introduction that is compatible with the conclusions of this study,
but that would have to be piloted in order to check its value, and its
relationship with better learning outcomes, as well as higher teacher
satisfaction.
Once that initial introduction of institutional messaging
automation tools is done, teachers (and students as well) will
probably prefer the kind of bots that alleviate bureaucratic or repetitive
tasks, such as answering frequent questions or answers on class or assignment
deadlines, as indicated by their answers to the respective
questions. These will help the introduction of more complex chatbots that will
affect more directly learning outcomes, such as chatbots that help students
integrate in the class, or are able to identify (and address) learning problems
in students or in groups of them. These should also be accompanied by
analytics on student interaction, as well as possibly some natural language
processing (in vernacular language) that will help assess the general mood of
the class, and how different material (or external factors) affect it.

% No remarkable differences between different groups.
This introduction of chatbots and chatbot technology will be helped by the fact
that no discernible differences are found between different groups. Even if
chatbots do have some potential for personalization or customization, based on
the class material, student and teaching staff, the general introduction of the
technology can be done in a general way and for all disciplines, teaching
experience and gender.

At any rate, the general feeling that transpires from the survey is that it is
essential for any institution to take into account the shareholder's opinion
when introducing chatbots. This is true almost across the board, but in the case
of chatbots (and instant messaging applications) their immediacy and the fact
that they can enter into what we could call the private sphere makes this even
more necessary.

One of the questions we made in the {\bf second survey}, related to
the messaging applications used by tertiary education teachers, opens a new line
of inquiry about what they {\em perceive} as such, and about how it is used.
Namely, the responses indicated that the teachers perceived as messaging
application not only what is usually called a chat or instant messaging app such
as WhatsApp or Microsoft Teams, but also the means provided by the university for
communicating with the student, such as a feature of grading applications that
will email the grade to the student. This implies that there is a need by
teachers, communication with students, also mostly unidirectional, and that it
does not matter so much how that need is covered. However, this will need careful
consideration, including how it ties with the automation of the learning
experience that the chatbots bring.

Our analysis of the results of this survey can be fed into several potential avenues for future research. For instance, the rollout of extensive videoconferencing and virtual
teaching solutions that the COVID pandemic has brought has also taught us a
series of lessons. It increases isolation, for instance, and decreases the
amount of synchronous contact that happens in the fringes of the classroom:
teaching staff offices, before and after class. A future line of work would be
focused on exactly this, and what kind of needs could be covered by chatbot
technology. We will create a series of international surveys that will
investigate this.

Finally, the full extent of chatbot technology is not really being examined in
these surveys. They can be connected to natural language processing engines with
sentiment analysis as well as other analytics. Examining the mood of the class,
as response to new material, assignment, exams or external events might help
will help any student-centered teaching strategy, which will help students (and
teachers) reach their learning objectives. This line of work, however, will be
based on the introduction of some pilot study, combined with initial opinion
assessment on the students.

\bibliographystyle{apalike}
\bibliography{edubots}

\begin{backmatter}

% ********************** DECLARATIONS ****************************

\section{Declarations}
\label{sec:decl}

\subsection{Availability of data and materials}

The datasets generated and analysed during the current study are available in the {\tt anonymous} repository, accessible from https://anonymous.URL
% {\tt https://github.com/JJ/edubots-paper/tree/main/data}.

\subsection{Competing interests}

The authors declare that they have no competing interests.

\subsection{Funding}

Suppressed for double blind review
% This work has been supported by EDUBOTS project, funded under the scheme Erasmus + KA2: Cooperation for innovation and the exchange of good practices - Knowledge Alliances (grant agreement no: 612446).

\subsection{Authors' contributions}

All the authors have contributed to the study presented in this manuscript. (Rest hidden for double blind)

% JJM has been the leader, drafted the first versions of the survey and supervised the writing of the manuscript.
%
% PAC has revised the paper and researched the state of the art, as well as contributed to the survey design.
% AMM has revised the paper and researched the state of the art, as well as contributed to the survey design.
% FB has made the analysis and processed data and contributed to the survey design.
% NA has supervised the writing  of the section devoted to the state of the art and written the bulk of it, as well as contributed to discussion and conclusions.
% OT has performed data analysis, and contributed to discussion and conclusions.
% AG has contributed to the design of the survey, analysis, discussion and conclusions.

All authors have read and approved the final manuscript.

\subsection{Acknowledgements}

% The authors want to thank to the educators who participated in the survey, and also people who have distributed the survey in their mailing lists, especially Mario García Valdez, Jesús Moreno León and the late José Raúl Canay.
We would also like to thank the anonymous reviewers that have contributed greatly to the improvement of this paper.

\section{Appendix I: Survey questions}
\label{app:questions}

The surveys have the same common initial demographic questions:\begin{itemize}
\item Degrees or titles where the teacher was working, either university or other tertiary education.
\item Discipline: Engineering, Social Sciences, Health and Bio, General Sciences, Humanities, Other.
\item Gender: Man, Woman, Rather not say.
\item Age: 25-35, 35-45, 45-55, more than 55.
\item Experience on the job: up to 5 years, 5-15 years, 15-25 years, more than 25 years.
\textcolor{black}{
\item Type of degree: university, other tertiary education. This question was implicit in the first survey: there were two different versions of the survey. However, it was an explicit question in the second survey.}
\end{itemize}

\begin{figure}[h!tbp]
\begin{center}
\includegraphics[width=0.85\textwidth]{figures/survey-1.png}
\end{center}
    \caption{Screen capture of the Spanish-language form used for the \bf{first survey}}
\label{figure:survey1}
\end{figure}
% Including questions in surveys
The {\bf first survey}, shown in its original form in Figure \ref{figure:survey1}, which \textcolor{black}{focused on} the use of messaging apps in class, includes the following questions:

\textit{1. Regarding the use of messaging apps in class, you prefer ...}
\begin{itemize}
\item A chat group with the students in the same course
\item Students self-organize and with the teacher not being part of any chat group
\textcolor{black}{
\item A chat group with all the students and teachers of the School/High-School/Faculty
}
\item A chat group with all the students and teachers of the same study year
\end{itemize}

\textit{2. Which messaging app platform do you use to communicate with your peers or students?}
\begin{itemize}
\item WhatsApp
\item Telegram
\item LinkedIn
\item Twitter messages
\item Discord
\item Slack
\item Snapchat
\item The platform provided by your own institution
\item None, I do not use any messaging app to communicate with my students
\end{itemize}

\textit{3. Which kind of chatbots (software that automatically respond to questions or commands) could be useful to improve the learning results with your students or help you managing your course?}
\begin{itemize}
\item Agenda bots that e.g. remind students about project deadlines
\item Bots that collect and provide answers to frequent questions
\item Bots that inform students about their grades
\item Bots that help with the class materials (e.g. searching for topics about a concept, or asking about them)
\end{itemize}

\textit{4. The virtualization of teaching due to the COVID pandemic during this year and the previous one, did it mean any change in your viewpoint or use of messaging apps in class?}
\begin{itemize}
\item No, I did use it before and use it now
\item Yes, I started using safer alternatives
\item Yes, I did not use any messaging platform before but I do use it now
\item No, I did not use any messaging platform before or now
\end{itemize}

\begin{figure}[h!tbp]
\begin{center}
\includegraphics[width=\textwidth]{figures/survey-2.png}
\end{center}
    \caption{Screen capture of the Spanish-language form used for the \bf{second survey}}
\label{figure:survey2}
\end{figure}
The {\bf second survey}, shown in its original form in Figure \ref{figure:survey2}, followed the same paths, only in this case there was a single questionnaire, and with a different scope, focusing on the kind of interaction media features that teachers value the most. %\textcolor{black}{A capture of these questions is included in Figure \ref{fig:survey2}} % this sentence is repeated
% Olia - see my comment above.
It was filled by teachers that were also students in a Professional Development course (on the use of new technologies in higher education, around 1/4 of them) as well as University of Granada teachers who knew about it by emails from the authors, or from the vicedeanship for International Relations that included it in its newsletter. This means that there might be a higher proportion of 1) teachers with few years of experience 2) teachers from the computer science faculty and 3) teachers from Granada. We think, however that there's no explicit bias in this selection, although of course specific percentages will vary.

Specifically, this survey asked the teachers, in addition to general information (such as gender, age, experience, etc), what are  their \textcolor{black}{past experiences and needs} with regard to the use of instant messaging applications with students, and even using bots or chatbots in the classroom. In addition the survey asked about the most relevant \textit{interaction media features} for the teachers with respect to the potential tool to use in class.

The specific questions included in this second survey are listed below:

% Antonio - I have included a reference name to each one of these variables, which are the names used on Tables 8 and 9. Thus it will be easy to recognize every variable, and then, we could refer to them in those tables.

\textit{1. You prefer the interaction with your students to be ...}
\begin{itemize}
\item Synchronous: students and teacher/s in the same space, simultaneously. [\texttt{Synchrony - Yes}]
\item Asynchronous: the student asks at anytime, the teacher responds whenever possible. [\texttt{Synchrony - No}]
\end{itemize}

\textit{2. Regarding the interaction with your students, you ...}
\begin{itemize}
\item Prefer it to be done by institutional means/tools: office hours, forum/messages via the virtual campus, institutional email. [\texttt{Officiality - Yes}]
\item Admit the use of other means/tools: blogs, chat groups in messaging platforms, any others. [\texttt{Officiality - No}]
\end{itemize}

\textit{3. Which media features do you value the most for your interaction with your students?}
\begin{itemize}
\item Familiarity with its use i.e. that I do not need to learn a new tool. [\texttt{Familiarity}]
\item The official formation/training provided by your institution on its use. [\texttt{Official formation}]
\item That only one-directional communication is allowed (from teachers to students). [\texttt{Unidirectionality}]
\item Non-hierarchical tools that allowed teacher-to-student and student-to-student communication. [\texttt{Horizontally}]
\item Connectivity to other tools, for instance common log in. [\texttt{Connectivity}]
\item Possibility for teachers to develop/add their own functionalities (e.g. automatic correction tools). [\texttt{Pluggability}]
\item Possibility for users to hide personal data (such as my telephone number). [\texttt{Hidden Phone Number}]
\item Not having to worry about maintenance and that it is easily sustainable from one year to the next. [\texttt{Sustainability}]
\item That it offers the chance to extract data to assess the student performance. [\texttt{Analytics}]
\end{itemize}

\end{backmatter}

\end{document}